---
title: "Interactive Maps with leaflet in R"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  pdf_document: default
  html_document:
    code_folding: hide
    fig_align: center
    fig_height: 6
    fig_width: 12
    keep_md: yes
editor_options:
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, leaflet, ggmap, leaflet.extras, htmltools, scales, htmlwidgets)

theme_set(theme_bw())
```

https://www.datacamp.com/courses/interactive-maps-with-leaflet-in-r

## Google Mapping API

https://cloud.google.com/maps-platform/

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location
register_google(key = "AIzaSyAfnLNZjvYdMx-cyga_qA1oJ6P36dRGalA") 
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## Setting up the map

Creating an Interactive Web Map
Similar to the packages in the tidyverse, the leaflet package makes use of the pipe operator (i.e., %>%) from the magrittr package to chain function calls together. This means we can pipe the result of one function into another without having to store the intermediate output in an object. For example, one way to find every car in the mtcars data set with a mpg >= 25 is to pipe the data through a series of functions.

mtcars  %>% 
    mutate(car = rownames(.))  %>% 
    select(car, mpg)  %>% 
    filter(mpg >= 25)  
To create a web map in R, you will chain together a series of function calls using the %>% operator. Our first function leaflet() will initialize the htmlwidget then we will add a map tile using the addTiles() function.

Instructions
100 XP
Load the leaflet library.
Call the leaflet() function.
Pipe the output of leaflet() into addTiles().
Experiment with zooming and panning your first interactive web map.

```{r ciwm}
# Load the leaflet library
library(leaflet)

# Create a leaflet map with default map tile using addTiles()
leaflet() %>%
    addTiles()
```

Provider Tiles
In the previous exercise, addTiles() added the default OpenStreetMap (OSM) tile to your leaflet map. Map tiles weave multiple map images together. The map tiles presented adjust when a user zooms or pans the map enabling the interactive features you experimented with in exercise 2.

The leaflet package comes with more than 100 map tiles that you can use. These tiles are stored in a list called providers and can be added to your map using addProviderTiles() instead of addTiles().

The leaflet and tidyverse libraries have been loaded for you.

Instructions
100 XP
Print the list of 100+ provider tiles that are included in the leaflet package.
To make this output more readable, print just the names of the provider tiles using the names() function.
Use the str_detect() function from the stringr package to determine which provider tiles include the string "CartoDB".
Print the name of every provider map tile that includes the string "CartoDB".

```{r PT}
# Print the providers list included in the leaflet library
providers

# Print only the names of the map tiles in the providers list 
names(providers)

# Use str_detect() to determine if the name of each provider tile contains the string "CartoDB"
str_detect(names(providers), "CartoDB")

# Use str_detect() to print only the provider tile names that include the string "CartoDB"
names(providers)[str_detect(names(providers), "CartoDB")]
```

Adding a Custom Map Tile
Did any tile names look familiar? If you have worked with the mapping software you may recognize the name ESRI or CartoDB.

We create our first leaflet map using the default OSM map tile.

leaflet() %>% 
    addTiles()
We will primarily use CartoDB provider tiles, but feel free to try others, like Esri. To add a custom provider tile to our map we will use the addProviderTiles() function. The first argument to addProviderTiles() is your leaflet map, which allows us to pipe leaflet() output directly into addProviderTiles(). The second argument is provider, which accepts any of the map tiles included in the providers list.

Familiarize yourself with the SCRIPT.R and HTML VIEWER tabs. Click back and forth to type your code and view your maps.

Instructions 1/3
Change addTiles() to addProviderTiles() and set the provider argument to "CartoDB".
2
Create a leaflet map that uses the "Esri" provider tile.
3
Create a leaflet map that uses the "CartoDB.PositronNoLabels" provider tile.

```{r acmt}
# Change addTiles() to addProviderTiles() and set the provider argument to "CartoDB"
leaflet() %>% 
    addProviderTiles(provider = "CartoDB")
# Create a leaflet map that uses the Esri provider tile 
leaflet() %>% 
    addProviderTiles("Esri")
# Create a leaflet map that uses the CartoDB.PositronNoLabels provider tile
leaflet() %>% 
    addProviderTiles("CartoDB.PositronNoLabels")
```

A Map with a View I
You may have noticed that, by default, maps are zoomed out to the farthest level. Rather than manually zooming and panning, we can load the map centered on a particular point using the setView() function.

leaflet()  %>% 
    addProviderTiles("CartoDB")  %>% 
    setView(lat = 40.7, lng = -74.0, zoom = 10)
Currently, DataCamp has offices at the following locations:

350 5th Ave, Floor 77, New York, NY 10118

Martelarenlaan 38, 3010 Kessel-Lo, Belgium

These addresses were converted to coordinates using the geocode() function in the ggmaps package.

NYC: (-73.98575, 40.74856)
Belgium: (4.717863, 50.881363)

Use the coordinates above to create a map with the "CartoDB" provider tile that is centered on DataCamp's New York office with a zoom of 6.

Instructions 2/2
50 XP
Now create a map with the "CartoDB.PositronNoLabels" provider tile that is centered on DataCamp's Belgium office with a zoom of 4.

```{r mwv}
# Map with CartoDB tile centered on DataCamp's NYC office with zoom of 6
leaflet()  %>% 
    addProviderTiles("CartoDB")  %>% 
    setView(lng = -73.98575, lat = 40.74856, zoom = 6)

dc_hq <- geocode(c("350 5th Ave, Floor 77, New York, NY 10118","Martelarenlaan 38, 3010 Kessel-Lo, Belgium"))

# Map with CartoDB.PositronNoLabels tile centered on DataCamp's Belgium office with zoom of 4
leaflet() %>% 
    addProviderTiles("CartoDB.PositronNoLabels") %>% 
    setView(lng = dc_hq$lon[2], lat = dc_hq$lat[2], zoom = 4)
```

A Map with a Narrower View
We can limit users' ability to pan away from the map's focus using the options argument in the leaflet() function. By setting minZoom anddragging, we can create an interactive web map that will always be focused on a specific area.

leaflet(options = 
        leafletOptions(minZoom = 14, dragging = FALSE))  %>% 
  addProviderTiles("CartoDB")  %>% 
  setView(lng = -73.98575, lat = 40.74856, zoom = 14) 
Alternatively, if we want our users to be able to drag the map while ensuring that they do not stray too far, we can set the maps maximum boundaries by specifying two diagonal corners of a rectangle.

You'll use dc_hq to create a map with the "CartoDB" provider tile that is centered on DataCamp's Belgium office.

Instructions
100 XP
Use a minimum zoom level of 12.
Set the dragging option to TRUE.
Use maximum bounds of .05 decimal degrees from the headquarters.

```{r mwnv}
leaflet(options = leafletOptions(
                    # Set minZoom and dragging 
                    minZoom = 12, dragging = TRUE))  %>% 
  addProviderTiles("CartoDB")  %>% 
  
  # Set default zoom level 
  setView(lng = dc_hq$lon[2], lat = dc_hq$lat[2], zoom = 14) %>% 
  
  # Set max bounds of map 
  setMaxBounds(lng1 = dc_hq$lon[2] + .05, 
               lat1 = dc_hq$lat[2] + .05, 
               lng2 = dc_hq$lon[2] - .05, 
               lat2 = dc_hq$lat[2] - .05) 
```

Mark it
So far we have been creating maps with a single layer: a base map. We can add layers to this base map similar to how you add layers to a plot in ggplot2. One of the most common layers to add to a leaflet map is location markers, which you can add by piping the result of addTiles() or addProviderTiles() into the add markers function.

For example, if we plot DataCamp's NYC HQ by passing the coordinates to addMarkers() as numeric vectors with one element, our web map will place a blue drop pin at the coordinate. In chapters 2 and 3, we will review some options for customizing these markers.

leaflet()  %>% 
    addProviderTiles("CartoDB")  %>% 
    addMarkers(lng = -73.98575, lat = 40.74856)
The dc_hq tibble is available in your work space.

```{r mi}
# Plot DataCamp's NYC HQ
leaflet() %>% 
    addProviderTiles("CartoDB") %>% 
    addMarkers(lng = dc_hq$lon[1], lat = dc_hq$lat[1])

# Plot DataCamp's NYC HQ with zoom of 12    
leaflet() %>% 
    addProviderTiles("CartoDB") %>% 
    addMarkers(lng = -73.98575, lat = 40.74856)  %>% 
    setView(lng = -73.98575, lat = 40.74856, zoom = 12)    

# Plot both DataCamp's NYC and Belgium locations
leaflet() %>% 
    addProviderTiles("CartoDB") %>% 
    addMarkers(lng = dc_hq$lon, lat = dc_hq$lat)
```

Adding Popups and Storing your Map
To make our map more informative we can add popups. To add popups that appear when a marker is clicked we need to specify the popup argument in the addMarkers() function. Once we have a map we would like to preserve, we can store it in an object. Then we can pipe this object into functions to add or edit the map's layers.

dc_nyc <- 
    leaflet() %>% 
        addTiles() %>% 
        addMarkers(lng = -73.98575, lat = 40.74856, 
                   popup = "DataCamp - NYC") 

dc_nyc %>% 
    setView(lng = -73.98575, lat = 40.74856, 
            zoom = 2)
Let's try adding popups to both DataCamp location markers and storing our map in an object.

```{r apasym}
# Store leaflet hq map in an object called map
 map <- leaflet() %>%
          addProviderTiles("CartoDB") %>%
          # Use dc_hq to add the hq column as popups
          addMarkers(lng = dc_hq$lon, lat = dc_hq$lat,
                     popup = dc_hq$hq)

# Center the view of map on the Belgium HQ with a zoom of 5 
map_zoom <- map %>%
      setView(lat = 50.881363, lng = 4.717863,
              zoom = 5)

# Print map_zoom
map_zoom
```

## Plotting Points

Cleaning up the Base Map
If you are storing leaflet maps in objects, there will come a time when you need to remove markers or reset the view. You can accomplish these tasks with the following functions.

clearMarkers()- Remove one or more features from a map
clearBounds()- Clear bounds and automatically determine bounds based on map elements
To remove the markers and to reset the bounds of our m map we would:

m <- m  %>% 
        addMarkers(lng = dc_hq$lon, lat = dc_hq$lat) %>% 
        setView(lat = 50.9, lng = 4.7, zoom = 5)

m  %>% 
    clearMarkers() %>% 
    clearBounds()
The leaflet map of DataCamp's headquarters has been printed for you.

```{r cubm}
# Remove markers, reset bounds, and store the updated map in the m object
map_clear <- map %>%
        clearMarkers() %>% 
        clearBounds()

# Print the cleared map
map_clear
```

Exploring the IPEDS Data II
Most analyses require data wrangling. Luckily, there are many functions in the tidyverse that facilitate data frame cleaning. For example, the drop_na() function will remove observations with missing values. By default, drop_na() will check all columns for missing values and will remove all observations with one or more missing values.

```{r eid2}
ipeds <- read.csv("ipeds.csv")

# Remove colleges with missing sector information
ipeds <- 
    ipeds %>% 
    drop_na()

# Count the number of four-year colleges in each state
ipeds %>% 
    group_by(state)  %>% 
    count()

# Create a list of US States in descending order by the number of colleges in each state
ipeds  %>% 
    group_by(state)  %>% 
    count()  %>% 
    arrange(desc(n))
```

California Colleges
Now it is your turn to map all of the colleges in a state. In this exercise, we'll apply our example of mapping Maine's colleges to California's colleges. The first step is to set up your data by filtering the ipeds data frame to include only colleges in California. 

```{r cc}
# Create a dataframe called `ca` with data on only colleges in California
ca <- ipeds %>%
        filter(state == "CA")

map <- leaflet() %>%
          addProviderTiles("CartoDB")

# Use `addMarkers` to plot all of the colleges in `ca` on the `m` leaflet map
map %>%
    addMarkers(lng = ca$lng, lat = ca$lat)
```

The City of Colleges
Based on our map of California colleges it appears that there is a cluster of colleges in and around the City of Angels (e.g., Los Angeles). Let's take a closer look at these institutions on our leaflet map.

The coordinates for the center of LA are provided for you in the la_coords data frame.

la_coords <- data.frame(lat = 34.05223, lon = -118.2437) 
Once you create a map focused on LA, try panning and zooming the map. Can you find the cluster of colleges East of LA known as the Claremont Colleges?

When there are hundreds of markers, do you find the pin markers helpful or do they get in your way?

The coordinates of LA have been provided in the la_coords data frame and the ca data frame of California colleges and the map have been loaded for you.

```{r tcoc}
la_coords <- data.frame(lat = 34.05223, lon = -118.2437)

# Set the zoom level to 8 and store in the m object
map_zoom <-
    map %>%
    addMarkers(data = ca) %>%
     setView(lng = la_coords$lon, lat = la_coords$lat, zoom = 8)

map_zoom

```

Circle Markers
Circle markers are notably different from pin markers:

We can control their size
They do not "stand-up" on the map
We can more easily change their color
There are many ways to customize circle markers and the design of your leaflet map. To get started we will focus on the following arguments.

addCircleMarkers(map, lng = NULL, lat = NULL, 
                 radius = 10, color = "#03F", popup = NULL)
The first argument map takes a leaflet object, which we will pipe directly into addCircleMarkers(). lng and lat are the coordinates we are mapping. The other arguments can customize the appearance and information presented by each marker.

The ca data frame and the leaflet object map have been loaded for you.

```{r cm}
# Clear the markers from the map 
map2 <- map %>% 
            clearMarkers()

# Change the radius of each circle to be 2 pixels and the color to red
map2 %>% 
    addCircleMarkers(lng = ca$lng, lat = ca$lat,
                     radius = 2, color = "red")
```

Making our Map Pop
Similar to building a plot with ggplot2 or manipulating data with dplyr, your map needs to be stored in an object if you reference it later in your code.

Speaking of dplyr, the %>% operator can pipe data into the function chain that creates a leaflet map.

ipeds %>% 
    leaflet()  %>% 
        addTiles() %>% 
        addCircleMarkers(popup = ~name, color = "#FF0000")

Piping makes our code more readable and allows us to refer to variables using the ~ operator rather than repeatedly specifying the data frame.

The color argument in addCircleMarkers() takes the name of a color or a hex code. For example, red or #FF0000.

map has been printed for you. Notice the circle markers are gone!

```{r momp}
# Add circle markers with popups for college names
map %>%
    addCircleMarkers(data = ca, radius = 2, popup = ~name)

# Change circle color to #2cb42c and store map in map_color object
map_color <- map %>% 
    addCircleMarkers(data = ca, radius = 2, color = "#2cb42c", popup = ~name)

# Print map_color
map_color
```

Building a Better Pop-up
With the paste0() function and a few html tags, we can customize our popups. paste0() converts its arguments to characters and combines them into a single string without separating the arguments.

addCircleMarkers(popup = ~paste0(name,
                                 "<br/>",
                                 sector_label))
We can use the <br/> tag to create a line break to have each element appear on a separate line.

To distinguish different data elements, we can make the name of each college italics by wrapping the name variable in <i></i>

addCircleMarkers(popup = ~paste0("<i>",
                                 name,
                                 "</i>", 
                                 "<br/>", 
                                 sector_label))
                                 
```{r bbpu}
# Clear the bounds and markers on the map object and store in map2
map2 <- map %>% 
            clearBounds() %>% 
            clearMarkers()

# Add circle markers with popups that display both the institution name and sector
map2 %>% 
    addCircleMarkers(data = ca, radius = 2, 
                     popup = ~paste0(name, "<br/>", sector_label))

# Make the institution name in each popup bold
map2 %>% 
    addCircleMarkers(data = ca, radius = 2, 
                     popup = ~paste0("<b>", name, "</b>", "<br/>", sector_label))
```

Swapping Popups for Labels
Popups are great, but they require a little extra effort. That is when labels come to our the aid. Using the label argument in the addCircleMarkers() function we can get more information about one of our markers with a simple hover!

ipeds %>% 
    leaflet()  %>% 
    addProviderTiles("CartoDB")  %>% 
    addCircleMarkers(label = ~name, radius = 2)
Labels are especially helpful when mapping more than a few locations as they provide quick access to detail about what each marker represents.

```{r spfl}
# Add circle markers with labels identifying the name of each college
map %>% 
    addCircleMarkers(data = ca, radius = 2, label = ~name)

# Use paste0 to add sector information to the label inside parentheses 
map %>% 
    addCircleMarkers(data = ca, radius = 2, label = ~paste0(name, " (", sector_label, ")"))
```

Creating a Color Palette using colorFactor
So far we have only used color to customize the style of our map. With colorFactor() we can create a color palette that maps colors the levels of a factor variable.

pal <- 
   colorFactor(palette = c("blue", "red", "green"), 
               levels = c("Public", "Private", "For-Profit"))

m %>% 
    addCircleMarkers(color = ~pal(sector_label))
Why might we not want to use this particular color palette?

If you are interested in using a continuous variable to color a map see colorNumeric().

pal <- colorNumeric(palette = "RdBu", domain = c(25:50))

ipeds %>% 
    leaflet() %>% 
        addProviderTiles("CartoDB")  %>% 
        addCircleMarkers(radius = 2, color = ~pal(lat))

```{r ccpucf}
# Make a color palette called pal for the values of `sector_label` using `colorFactor()`  
# Colors should be: "red", "blue", and "#9b4a11" for "Public", "Private", and "For-Profit" colleges, respectively
pal <- colorFactor(palette = c("red", "blue", "#9b4a11"), 
                   levels = c("Public", "Private", "For-Profit"))

# Add circle markers that color colleges using pal() and the values of sector_label
map2 <- 
    map %>% 
        addCircleMarkers(data = ca, radius = 2, 
                         color = ~pal(sector_label), 
                         label = ~paste0(name, " (", sector_label, ")"))

# Print map2
m <- map2
map2
```

A Legendary Map
Adding information to our map using color is great, but it is only helpful if we remember what the colors represent. With addLegend() we can add a legend to remind us.

There are several arguments we can use to custom the legend to our liking, including opacity, title, and position. To create a legend for our colorNumeric() example, we would do the following.

```{r alm}
# Customize the legend
map2 %>% 
    addLegend(pal = pal, 
              values = c("Public", "Private", "For-Profit"),
              # opacity of .5, title of Sector, and position of topright
              opacity = 0.5, title = "Sector", position = "topright")
```

## Layers

Middle America
Starting from scratch, our first step is to create a base map and to set our map's view. Rather than using the geocode() function from ggmaps, we'll add an extra feature to our map so we can use the map to geocode.

After creating the map in Step 1, search for "U.S. Center Chapel " in Kansas and click the marker to add a pin.

When you hover over the pin what are the coordinates?

```{r ma}
library(leaflet.extras)

leaflet() %>%
  addTiles() %>% 
  addSearchOSM() %>% 
  addReverseSearchOSM() 
```

Building a Base
Now that we know where to center our map (Lat = 39.8282, Lng = -98.5795), let's build a new basemap. Remember that we can use either addTiles() or addProviderTiles() to add a map tile to our leaflet object.

Since we are working toward mapping all of America's colleges, we can include the ipeds data at the start of the chain of functions that will create our base map. This way the ipeds data will be stored with our base map so that we can easily access it. Once we have built a solid base, we can map every college in America with just two lines of R code.

The ipeds data and leaflet and tidyverse libraries have been loaded for you.

```{r bab}
m2 <- 
    ipeds %>% 
        leaflet() %>% 
            # use the CartoDB provider tile
            addProviderTiles("CartoDB") %>% 
            # center on the middle of the US with zoom of 3
            setView(lat = 39.8282, lng = -98.5795, zoom = 3)

# Map all American colleges 
m2 %>% 
    addCircleMarkers(lng = ~lng, lat = ~lat,
                     radius = 1, color = "red") 
```

Mapping Public Colleges
We can add each sector to our map as a layer providing users with the ability to select which sectors are displayed. To do this we will make use of a new argument to the addCircleMarkers() function, called a group.

leaflet() %>% 
        addProviderTiles("CartoDB") %>% 
        addCircleMarkers(data = public, 
                         label = ~htmlEscape(name), 
                         group = "Public")
We'll integrate another feature into our map from the htmltools library. Wrapping our labels with the htmlEscape() function will sanitize characters that may be interpreted as HTML. This will prevent any of the college names from appearing with unintended formatting.

The color palette pal we created in Chapter 2 has been loaded for you.

```{r mpc}
# Load the htmltools package
library(htmltools)

# Create data frame called public with only public colleges
public <- subset(ipeds, sector_label == "Public")  

# Create a leaflet map of public colleges called m3 
m3 <- leaflet() %>% 
        addProviderTiles("CartoDB") %>% 
        addCircleMarkers(data = public, radius = 1, label = ~htmlEscape(name),
                         color = ~pal(sector_label), group = "Public")

m3
```

Mapping Public and Private Colleges
We can add private colleges exactly how we added public colleges. Then using the addLayersControl() function with the overlayGroups argument we can give our users the ability to display public and/or private colleges. The overlayGroups argument takes a vector of groups that we have defined while creating our layers (i.e., public and private). As a reminder, here is how we can add public colleges to our map as a layer:

# Create data frame called public with only public colleges
public <- filter(ipeds, sector_label == "Public")  

# Add public colleges as a layer and save map as `m3`
m3 <- leaflet() %>% 
        addProviderTiles("CartoDB") %>% 
        addCircleMarkers(data = public, radius = 2, 
                         label = ~htmlEscape(name),
                         color = ~pal(sector_label),
                         group = "Public") %>% 
        addLayersControl(overlayGroups = c("Public"))
The htmltools library, color palette pal and the m3 map with public colleges have been loaded for you.

```{r mppc}
# Create data frame called private with only private colleges
private <- filter(ipeds, sector_label == "Private")  

# Add private colleges to `m3` as a new layer
m3 <- m3 %>% 
        addCircleMarkers(data = private, radius = .5, label = ~htmlEscape(name),
                         color = ~pal(sector_label), group = "Private") %>% 
        addLayersControl(overlayGroups = c("Public", "Private"))

m3
```

Mapping All Colleges
Let's keep layering up and add for-profit colleges to our leaflet map that is stored in the m3 object.

After you print() the m3 map with public, private, and for-profit colleges as their own layers, then try removing all three layers and adding them back to the map in a different order.

The different college sectors are added back to the map as layers in the order you specify (i.e., the last sector that you select will be on top).

The color palette pal and the m3 map with public and private colleges have been loaded for you.

```{r mac}
# Create data frame called profit with only For-Profit colleges
profit <- filter(ipeds, sector_label == "For-Profit")  

# Add For-Profit colleges to `m3` as a new layer
m3 <- m3 %>% 
        addCircleMarkers(data = profit, radius = 2, label = ~htmlEscape(name),
                         color = ~pal(sector_label),   group = "For-Profit")  %>% 
        addLayersControl(overlayGroups = c("Public", "Private", "For-Profit"))  

# Center the map on the middle of the US with a zoom of 4
m4 <- m3 %>%
        setView(lat = 39.8282, lng = -98.5795, zoom = 4) 
        
m4
```

Change up the Base
Similar to how we added overlay groups (i.e., college sectors), we can allow our users to toggle between different base maps using the baseGroups argument to the addLayersControl() function.

First we will need to add multiple basemaps and specify the name of a group for each.

leaflet() %>% 
  addTiles(group = "OSM") %>% 
  addProviderTiles("CartoDB", group = "CartoDB") 
Try running the above code in the console. You should see a leaflet map in the viewer with the CartoDB basemap. This is because we added the CartoDB base map after the default OpenStreetMap tile. If we add addLayerControl(), our users will be able to toggle between the two base maps (you can include many base maps, but only one can be selected at a time).

```{r cutb}
leaflet() %>% 
  # Add the OSM, CartoDB and Esri tiles
  addTiles(group = "OSM") %>% 
  addProviderTiles("CartoDB", group = "CartoDB") %>% 
  addProviderTiles("Esri", group = "Esri") %>% 
  # Use addLayersControl to allow users to toggle between basemaps
  addLayersControl(baseGroups = c("OSM", "CartoDB", "Esri"))
```

Putting it all Together
Let's try putting this all together at one time. Building our interactive map of every four-year college in America can be broken down into four steps:

Initialize the web map
Lay down our base maps
Plot each sector of colleges as its own layer
Add layer controls so users can toggle base and overlay groups
The public, private, and profit data frames as well as the color palette pal have been loaded for you.

```{r piat}
m4 <- leaflet() %>% 
        addTiles(group = "OSM") %>% 
        addProviderTiles("CartoDB", group = "Carto") %>% 
        addProviderTiles("Esri", group = "Esri") %>% 
        addCircleMarkers(data = public, radius = 2, label = ~htmlEscape(name),
                         color = ~pal(sector_label),  group = "Public") %>% 
        addCircleMarkers(data = private, radius = 2, label = ~htmlEscape(name),
                           color = ~pal(sector_label), group = "Private")  %>% 
        addCircleMarkers(data = profit, radius = 2, label = ~htmlEscape(name),
                         color = ~pal(sector_label), group = "For-Profit")  %>% 
        addLayersControl(baseGroups = c("OSM", "Carto", "Esri"), 
                         overlayGroups = c("Public", "Private", "For-Profit")) %>% 
        setView(lat = 39.8282, lng = -98.5795, zoom = 4) 

m4
```

Adding a Piece of Flair
Earlier in this chapter, we used addSearchOSM() to find the middle of the US. To search for markers, rather than locations, we can use the addSearchFeatures() function. addSearchFeatures() will add a search box that you can use to find markers in the group(s) passed to the targetGroups argument. For example, to search our map for public colleges we could use the following code.

m4  %>% 
  addSearchFeatures(
    targetGroups = 'Public',
    options = searchFeaturesOptions(zoom = 10)) 
The m4 map and the leaflet.extras package have been loaded for you.

```{r aipof}
# Make each sector of colleges searchable 
m4_search <- m4  %>% 
  addSearchFeatures(
    targetGroups = c("Public", "Private", "For-Profit"), 
    # Set the search zoom level to 18
    options = searchFeaturesOptions(zoom = 18)) 

# Try searching the map for a college
m4_search
```

A Cluster Approach
Rather than using layers to improve the usability of our map, we could elect to cluster the colleges by clustering groups of nearby colleges together to reduce the number of points on the map. Zooming in will cause the clusters to break apart and the individual colleges to appear. This can be a useful tactic for mapping a large number of points in a user-friendly manner.

We can cluster all of our colleges by setting the clusterOptions argument of addCircleMarkers() as follows.

ipeds %>% 
  leaflet() %>%
    addTiles() %>% 
    addCircleMarkers(clusterOptions = markerClusterOptions())
The ipeds data, htmltools library, and color palette pal have been loaded for you.

```{r aca}
ipeds %>% 
    leaflet() %>% 
        addTiles() %>% 
        # Sanitize any html in our labels
        addCircleMarkers(radius = 2, label = ~htmlEscape(name),
                         # Color code colleges by sector using the `pal` color palette
                         color = ~pal(sector_label),
                         # Cluster all colleges using `clusterOptions`
                         clusterOptions = markerClusterOptions()) 
```

## Polygons

Introduction to Spatial Data
We have been mapping points, but there are several spatial features that can be mapped, including polygons. In R, polygons are often stored in a SpatialPolygonsDataFrame that holds the polygon, coordinate information, and a data frame with one row per polygon.

A SpatialPolygonsDataFrame called shp that contains the zip code boundaries for North Carolina has been loaded for you. shp has five slots that store various types of information:

data: data associated with each polygon
polygons: coordinates to plot polygons
plotOrder: order in which polygons are plotted
bbox: bounding box for geographic data (i.e., a rectangle)
proj4string: coordinate reference system
Let's take a closer look inside the shp object.

```{r itsd}
load("nc_zips.rda")
summary(shp)
class(shp)
slotNames(shp)
glimpse(shp@data)
class(shp@data)
shp@data$GEOID10
nc_income <- read.csv("nc_income.csv")
nc_income <- as.data.frame(nc_income)
nc_income$zipcode<-as.factor(nc_income$zipcode)
```

Joining Spatial Data
We can join data onto the data frame stored in the data slot of our SpatialPolygonsDataFrame. In this chapter, we are interested in the mean income at the zip code level as reported by the IRS. Once we have the income data joined onto the information in the data slot of shp we can map the mean income of zip codes on our leaflet map.

A data frame called nc_income has been loaded for you. Let's get started by taking a look at the nc_income data. Then we'll join it onto the information in the data slot of shp and see if there are any zip codes in our data that are missing income information.

```{r jsd}
# Glimpse the nc_income data
glimpse(nc_income)

# Summarize the nc_income data
summary(nc_income)

# Left join nc_income onto shp@data and store in shp_nc_income
shp_nc_income <- shp@data %>% 
                left_join(nc_income, by = c("GEOID10" = "zipcode"))

# Print the number of missing values of each variable in shp_nc_income
shp_nc_income  %>%
  summarize_all(funs(sum(is.na(.))))
```

addPolygons() Function
Let's look at those zip codes with missing data to hypothesize why they do not have income data.

We are mapping ZCTAs (not actual zip codes) so not every part of NC will have a boundary. Our boundaries may overlap because the file was simplified to reduce size. These are trade offs to consider when mapping polygons.

Enough nuance, let's make a map. You can pipe the shp data directly into our calls to leaflet(), addTiles(), and addPolygons() without supplying any additional arguments to map North Carolina's zip codes. To get you started, the shp SpatialPolygonsDataFrame including the IRS income variables has been loaded for you.

```{r apf}
# map the polygons in shp
nc_income <- as_tibble(nc_income)
shp2 <- shp
shp2@data <- as_tibble(shp2@data)
shp2@data <- shp2@data %>% rename(zipcode = GEOID10)
shp2@data <- left_join(shp2@data,nc_income)

shp <- shp2

shp %>% 
    leaflet() %>% 
    addTiles() %>% 
    addPolygons()

# which zips were not in the income data?
shp_na <- shp[is.na(shp$mean_income),]

# map the polygons in shp_na
shp_na %>% 
    leaflet() %>% 
    addTiles() %>% 
    addPolygons()
```

NC High Income Zips
Did you have a hypothesis of why certain zip codes are missing income information? It looks to me like many of them are areas that likely have low populations (e.g., parks, colleges, etc.) and the IRS only reports income data on zip codes with more than 100 filers.

Now let's focus in on a subset of zip codes with income data, namely the 25% of zip codes in NC with the highest mean incomes. Where do think these will fall within the states?

Let's take a look and find out.

```{r NChiz}
# summarize the mean income variable
summary(shp$mean_income)

# subset shp to include only zip codes in the top quartile of mean income
high_inc <- shp[!is.na(shp@data$mean_income) & shp$mean_income > 55917,]

# map the boundaries of the zip codes in the top quartile of mean income
high_inc %>%
  leaflet() %>%
  addProviderTiles("OpenTopoMap") %>% 
  addPolygons()
```

addPolygon() Options
So far we have used the default appearance for addPolygons(). There are several more ways to customize the polygons.

The arguments to addPolygons() we will focus on are:

weight: the thickness of the boundary lines in pixels
color: the color of the polygons
label: the information to appear on hover
highlightOptions: options to highlight a polygon on hover
addPolygons(weight = 2,
           color = "red",
           label = ~paste0("Total Income: ", dollar(income)),
           highlight = highlightOptions(weight = 10,
                                       color = "blue",
                                       bringToFront = TRUE))
The high_inc SpatialPolygonsDataFrame you created in the previous exercise has been loaded for you.

```{r apo}
# create color palette with colorNumeric()
nc_pal <- colorNumeric("YlGn", domain = high_inc@data$mean_income)

high_inc %>%
  leaflet() %>%
  addTiles() %>%
  # set boundary thickness to 1 and color polygons
  addPolygons(weight = 1, color = ~nc_pal(mean_income),
              # add labels that display mean income
              label = ~paste0("Mean Income: ", dollar(mean_income)),
              # highlight polygons on hover
              highlight = highlightOptions(weight = 5, color = "white",
              bringToFront = TRUE))
```

Let's do some Logging
Let's look at the quartiles of mean zip code incomes in NC.

summary(high_inc$mean_income)

   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  55947   62380   71655   83682   90695  550849
The range is nearly $500,000! However, the median is much closer to the min than to the max, indicating a right-skew. Since the mean income variable contains exceptionally large values, the continuous color gradient is not very helpful. Log transforming a right-skewed variable pulls large values closer to the mean and yields a more symmetrically distributed variable.

Log transforming the mean income on our map increases the variation in the color gradient across the high income zip codes and enables better visualization of the distribution of mean income across the state.

```{r ldsl}
# Use the log function to create a new version of nc_pal
nc_pal <- colorNumeric("YlGn", domain = log(high_inc@data$mean_income))

# comment out the map tile
high_inc %>%
  leaflet() %>%
  addProviderTiles("CartoDB") %>%
  # apply the new nc_pal to the map
  addPolygons(weight = 1, color = ~nc_pal(log(mean_income)), fillOpacity = 1,
              label = ~paste0("Mean Income: ", dollar(mean_income)),
              highlightOptions = highlightOptions(weight = 5, color = "white", bringToFront = TRUE))
```

Wealthiest Zip Codes in America
Zooming out from North Carolina, let's put together a polygon layer for the entire country. Then we will work to add this layer to our map of American colleges.

According to the 2015 IRS data, there are 427 zip codes in America with a mean income of $200,000 or more. It might be interesting to know which colleges are located in these affluent areas. Let's map these 427 polygons and explore our leaflet map to better understand where the highest income zip codes in America are. To get us started, a SpatialPolygonsDataFrame called wealthy_zip that contains information on each zip code in America with a mean income of $200,000 or more has been loaded for you.

Before we work with this data, take a minute to hypothesize where many of these zip codes are located.

```{r wzia}
#load("wealthiest_zips.Rda")

# plot zip codes with mean incomes >= $200k
wealthy_zips %>% 
  leaflet() %>% 
  addProviderTiles("CartoDB") %>% 
  # set color to green and create Wealth Zipcodes group
  addPolygons(weight = 1, fillOpacity = .7, color = "green",  group = "Wealthy Zipcodes", 
              label = ~paste0("Mean Income: ", dollar(mean_income)),
              highlightOptions = highlightOptions(weight = 5, color = "white", bringToFront = TRUE))
```

Final Map
Because our leaflet map is built in layers, we can add different types of information to the same base map (e.g., points and polygons). When adding new layers based on different data to an existing leaflet object we must specify the data argument with the function that creates the new layer to override the data that is piped through the chain of functions.

m4  %>% 
    addCircleMarkers(data = private)
Let's combine our point and polygon maps to add a layer highlighting America's wealthiest zip codes to our map of every college in America. To get us started, the last layered version of the college map m4 has been printed for you and the wealthy_zips SpatialPolygonsDataFrame is pre-loaded.

After you create the final map, take a few minutes to explore (and enjoy) your leaflet map!

```{r}
# Add polygons using wealthy_zips
final_map <- m4 %>% 
   addPolygons(data = wealthy_zips, weight = 1, fillOpacity = .5, color = "Grey",  group = "Wealthy Zip Codes", 
              label = ~paste0("Mean Income: ", dollar(mean_income)),
              highlightOptions = highlightOptions(weight = 5, color = "white", bringToFront = TRUE)) %>% 
    # Update layer controls including "Wealthy Zip Codes"
    addLayersControl(baseGroups = c("OSM", "Carto", "Esri"), 
                         overlayGroups = c("Public", "Private", "For-Profit", "Wealthy Zip Codes"))     

# Print and explore your very last map of the course!
final_map
```



