---
title: "Machine Learning Toolbox"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(caret)
library(MASS) #for the Boston dataset
library(mlbench) # for the Sonar dataset
library(e1071) # for confusionmatrix
library(caTools) # for the ROC curve
library(ranger) # for the ranger package that is an upgrade of randomForest
library(glmnet)
library(RANN) # to use knnImoute function
select <- dplyr::select # MASS also has a select function
library(caretEnsemble)
```

## Background

https://campus.datacamp.com/courses/machine-learning-toolbox/regression-models-fitting-them-and-evaluating-their-performance?ex=1#

http://topepo.github.io/caret/index.html

## Regression models:

In-sample error


```{r ISE}
# Fit a model to the mtcars data
data(mtcars)
model <- lm(mpg ~ hp, mtcars[1:20, ])

# Predict in-sample
predicted <- predict(model, mtcars[1:20, ], type = "response")

# Calculate RMSE
actual <- mtcars[1:20, "mpg"]
sqrt(mean((predicted - actual) ^ 2))
```

In-sample RMSE for linear regression on diamonds
As you saw in the video, included in the course is the diamonds dataset, which is a classic dataset from the ggplot2 package. The dataset contains physical attributes of diamonds as well as the price they sold for. One interesting modeling challenge is predicting diamond price based on their attributes using something like a linear regression.

Recall that to fit a linear regression, you use the lm() function in the following format:

mod <- lm(y ~ x, my_data)
To make predictions using mod on the original data, you call the predict() function:

pred <- predict(mod, my_data)

```{r isrmse}
# Fit lm model: model
data("diamonds")
model <- lm(price ~ ., diamonds)

# Predict on full data: p
p <- predict(model, diamonds, type = "response")

# Compute errors: error
actual <- diamonds$price
error <- p - actual

# Calculate RMSE
sqrt(mean((p - actual) ^ 2))
```

Out-of-sample error

```{r oose}
# Fit a model to the mtcars data
data(mtcars)
model <- lm(mpg ~ hp, mtcars[1:20, ])

# Predict in-sample
predicted <- predict(model, mtcars[21:32, ], type = "response")

# Calculate RMSE
actual <- mtcars[21:32, "mpg"]
sqrt(mean((predicted - actual) ^ 2))
```

Randomly order the data frame
One way you can take a train/test split of a dataset is to order the dataset randomly, then divide it into the two sets. This ensures that the training set and test set are both random samples and that any biases in the ordering of the dataset (e.g. if it had originally been ordered by price or size) are not retained in the samples we take for training and testing your models. You can think of this like shuffling a brand new deck of playing cards before dealing hands.

First, you set a random seed so that your work is reproducible and you get the same random split each time you run your script:

set.seed(42)
Next, you use the sample() function to shuffle the row indices of the diamonds dataset. You can later use these indices to reorder the dataset.

rows <- sample(nrow(diamonds))
Finally, you can use this random vector to reorder the diamonds dataset:

diamonds <- diamonds[rows, ]

```{r scramble}
# Set seed
set.seed(42)

# Shuffle row indices: rows
rows <- sample(nrow(diamonds))

# Randomly order data
shuffled_diamonds <- diamonds[rows, ]
```

Try an 80/20 split
Now that your dataset is randomly ordered, you can split the first 80% of it into a training set, and the last 20% into a test set. You can do this by choosing a split point approximately 80% of the way through your data:

split <- round(nrow(mydata) * .80)
You can then use this point to break off the first 80% of the dataset as a training set:

mydata[1:split, ]
And then you can use that same point to determine the test set:

mydata[(split + 1):nrow(mydata), ]

```{r split}
# Determine row to split on: split
split <- round(nrow(shuffled_diamonds) * .80)

# Create train
train <- shuffled_diamonds[1:split, ]

# Create test
test <- shuffled_diamonds[(split + 1):nrow(diamonds), ]
```

Predict on test set
Now that you have a randomly split training set and test set, you can use the lm() function as you did in the first exercise to fit a model to your training set, rather than the entire dataset. Recall that you can use the formula interface to the linear regression function to fit a model with a specified target variable using all other variables in the dataset as predictors:

mod <- lm(y ~ ., training_data)
You can use the predict() function to make predictions from that model on new data. The new dataset must have all of the columns from the training data, but they can be in a different order with different values. Here, rather than re-predicting on the training set, you can predict on the test set, which you did not use for training the model. This will allow you to determine the out-of-sample error for the model in the next exercise:

p <- predict(model, new_data)

```{r pots}
# Fit lm model on train: model
model <- lm(price ~ ., train)

# Predict on test: p
p <- predict(model, test)
```

Calculate test set RMSE by hand
Now that you have predictions on the test set, you can use these predictions to calculate an error metric (in this case RMSE) on the test set and see how the model performs out-of-sample, rather than in-sample as you did in the first exercise. You first do this by calculating the errors between the predicted diamond prices and the actual diamond prices by subtracting the predictions from the actual values.

Once you have an error vector, calculating RMSE is as simple as squaring it, taking the mean, then taking the square root:

sqrt(mean(error^2))

```{r CRMSE}
# Compute errors: error
error <- p - test[["price"]]

# Calculate RMSE
sqrt(mean(error^2))

# Compute errors: error
error <- p - test$price

# Calculate RMSE
sqrt(mean(error^2))
```

10-fold cross-validation
As you saw in the video, a better approach to validating models is to use multiple systematic test sets, rather than a single random train/test split. Fortunately, the caret package makes this very easy to do:

model <- train(y ~ ., my_data)
caret supports many types of cross-validation, and you can specify which type of cross-validation and the number of cross-validation folds with the trainControl() function, which you pass to the trControl argument in train():

model <- train(
  y ~ ., 
  my_data,
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 10,
    verboseIter = TRUE
  )
)
It's important to note that you pass the method for modeling to the main train() function and the method for cross-validation to the trainControl() function.

```{r Tfcv}
# Fit lm model using 10-fold CV: model
model <- train(price ~ .,diamonds, method = "lm", trControl = trainControl(method = "cv",number = 10, verboseIter = TRUE))

# Print model to console
model
```

5-fold cross-validation
In this course, you will use a wide variety of datasets to explore the full flexibility of the caret package. Here, you will use the famous Boston housing dataset, where the goal is to predict median home values in various Boston suburbs.

You can use exactly the same code as in the previous exercise, but change the dataset used by the model:

model <- train(
  medv ~ ., 
  Boston, # <- new!
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 10,
    verboseIter = TRUE
  )
)
Next, you can reduce the number of cross-validation folds from 10 to 5 using the number argument to the trainControl() argument:

trControl = trainControl(
  method = "cv", 
  number = 5,
  verboseIter = TRUE
)

```{r ffcv}
# Fit lm model using 10-fold CV: model
model <- train(medv ~ .,Boston, method = "lm", trControl = trainControl(method = "cv",number = 5, verboseIter = TRUE))

# Print model to console
model
```

5 x 5-fold cross-validation
You can do more than just one iteration of cross-validation. Repeated cross-validation gives you a better estimate of the test-set error. You can also repeat the entire cross-validation procedure. This takes longer, but gives you many more out-of-sample datasets to look at and much more precise assessments of how well the model performs.

One of the awesome things about the train() function in caret is how easy it is to run very different models or methods of cross-validation just by tweaking a few simple arguments to the function call. For example, you could repeat your entire cross-validation procedure 5 times for greater confidence in your estimates of the model's out-of-sample accuracy, e.g.:

trControl = trainControl(
  method = "repeatedcv", 
  number = 5,
  repeats = 5, 
  verboseIter = TRUE
)

```{r fbffcv}
# Fit lm model using 5 x 5-fold CV: model
model <- train(
  medv ~ ., 
  Boston,
  method = "lm",
  trControl = trainControl(
    method = "repeatedcv", 
    number = 5,
    repeats = 5, 
    verboseIter = TRUE
  )
)

# Print model to console
model
```

Making predictions on new data
Finally, the model you fit with the train() function has the exact same predict() interface as the linear regression models you fit earlier in this chapter.

After fitting a model with train(), you can simply call predict() with new data, e.g:

predict(my_model, new_data)

```{r mpnd}
# Predict on full Boston dataset
predict(model, Boston)
```

## Classification Models

Try a 60/40 split
As you saw in the video, you'll be working with the Sonar dataset in this chapter, using a 60% training set and a 40% test set. We'll practice making a train/test split one more time, just to be sure you have the hang of it. Recall that you can use the sample() function to get a random permutation of the row indices in a dataset, to use when making train/test splits, e.g.:

n_obs <- nrow(my_data)
permuted_rows <- sample(n_obs)
And then use those row indices to randomly reorder the dataset, e.g.:

my_data <- my_data[permuted_rows, ]
Once your dataset is randomly ordered, you can split off the first 60% as a training set and the last 40% as a test set.

```{r t60/40 split}
data(Sonar)

# Get the number of observations
n_obs <- nrow(Sonar)

# Shuffle row indices: permuted_rows
permuted_rows <- sample(n_obs)

# Randomly order data: Sonar
Sonar_shuffled <- Sonar[permuted_rows, ]

# Identify row to split on: split
split <- round(n_obs * 0.6)

# Create train
train <- Sonar_shuffled[1:split, ]

# Create test
test <- Sonar_shuffled[(split + 1):n_obs, ]

nrow(train)/nrow(Sonar)
```

Fit a logistic regression model
Once you have your random training and test sets you can fit a logistic regression model to your training set using the glm() function. glm() is a more advanced version of lm() that allows for more varied types of regression models, aside from plain vanilla ordinary least squares regression.

Be sure to pass the argument family = "binomial" to glm() to specify that you want to do logistic (rather than linear) regression. For example:

glm(Target ~ ., family = "binomial", dataset)
Don't worry about warnings like glm.fit: algorithm did not converge or glm.fit: fitted probabilities numerically 0 or 1 occurred. These are common on smaller datasets and usually don't cause any issues. They typically mean your dataset is perfectly separable, which can cause problems for the math behind the model, but R's glm() function is almost always robust enough to handle this case with no problems.

Once you have a glm() model fit to your dataset, you can predict the outcome (e.g. rock or mine) on the test set using the predict() function with the argument type = "response":

predict(my_model, test, type = "response")

```{r flrm}
# Fit glm model: model
model <- glm(Class ~ ., family = "binomial", train)

# Predict on test: p
p <- predict(model, test, type = "response")
```

Calculate a confusion matrix
As you saw in the video, a confusion matrix is a very useful tool for calibrating the output of a model and examining all possible outcomes of your predictions (true positive, true negative, false positive, false negative).

Before you make your confusion matrix, you need to "cut" your predicted probabilities at a given threshold to turn probabilities into a factor of class predictions. Combine ifelse() with factor() as follows:

pos_or_neg <- ifelse(probability_prediction > threshold, positive_class, negative_class)
p_class <- factor(pos_or_neg, levels = levels(test_values))
confusionMatrix() in caret improves on table() from base R by adding lots of useful ancillary statistics in addition to the base rates in the table. You can calculate the confusion matrix (and the associated statistics) using the predicted outcomes as well as the actual outcomes, e.g.:

confusionMatrix(p_class, test_values)

```{r ccm}
# If p exceeds threshold of 0.5, M else R: m_or_r
m_or_r <- ifelse(p > .5, "M", "R")

# Convert to factor: p_class
p_class <- factor(m_or_r, levels = levels(test[["Class"]]))

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```

Try another threshold
In the previous exercises, you used a threshold of 0.50 to cut your predicted probabilities to make class predictions (rock vs mine). However, this classification threshold does not always align with the goals for a given modeling problem.

For example, pretend you want to identify the objects you are really certain are mines. In this case, you might want to use a probability threshold of 0.90 to get fewer predicted mines, but with greater confidence in each prediction.

The code pattern for cutting probabilities into predicted classes, then calculating a confusion matrix, was shown in Exercise 7 of this chapter.

```{r tat}
# If p exceeds threshold of 0.9, M else R: m_or_r
m_or_r <- ifelse(p > .9, "M", "R")

# Convert to factor: p_class
p_class <- factor(m_or_r, levels = levels(test[["Class"]]))

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```

```{r fptcm}
# If p exceeds threshold of 0.9, M else R: m_or_r
m_or_r <- ifelse(p > .1, "M", "R")

# Convert to factor: p_class
p_class <- factor(m_or_r, levels = levels(test[["Class"]]))

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```

Plot an ROC curve
As you saw in the video, an ROC curve is a really useful shortcut for summarizing the performance of a classifier over all possible thresholds. This saves you a lot of tedious work computing class predictions for many different thresholds and examining the confusion matrix for each.

My favorite package for computing ROC curves is caTools, which contains a function called colAUC(). This function is very user-friendly and can actually calculate ROC curves for multiple predictors at once. In this case, you only need to calculate the ROC curve for one predictor, e.g.:

colAUC(predicted_probabilities, actual, plotROC = TRUE)
The function will return a score called AUC (more on that later) and the plotROC = TRUE argument will return the plot of the ROC curve for visual inspection.

```{r PROCC}
# Predict on test: p
p <- predict(model, test, type = "response")

# Make ROC curve
colAUC(p, test$Class, plotROC = TRUE)
```

Customizing trainControl
As you saw in the video, area under the ROC curve is a very useful, single-number summary of a model's ability to discriminate the positive from the negative class (e.g. mines from rocks). An AUC of 0.5 is no better than random guessing, an AUC of 1.0 is a perfectly predictive model, and an AUC of 0.0 is perfectly anti-predictive (which rarely happens).

This is often a much more useful metric than simply ranking models by their accuracy at a set threshold, as different models might require different calibration steps (looking at a confusion matrix at each step) to find the optimal classification threshold for that model.

You can use the trainControl() function in caret to use AUC (instead of acccuracy), to tune the parameters of your models. The twoClassSummary() convenience function allows you to do this easily.

When using twoClassSummary(), be sure to always include the argument classProbs = TRUE or your model will throw an error! (You cannot calculate AUC with just class predictions. You need to have class probabilities as well.)

```{r CtC}
# Create trainControl object: myControl
myControl <- trainControl(
  method = "cv",
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE
)
```

Using custom trainControl
Now that you have a custom trainControl object, it's easy to fit caret models that use AUC rather than accuracy to tune and evaluate the model. You can just pass your custom trainControl object to the train() function via the trControl argument, e.g.:

train(<standard arguments here>, trControl = myControl)
This syntax gives you a convenient way to store a lot of custom modeling parameters and then use them across multiple different calls to train(). You will make extensive use of this trick in Chapter 5.

```{r UctC, warning=FALSE}
# Train glm with custom trainControl: model
model <- train(
  Class ~ ., 
  Sonar, 
  method = "glm",
  trControl = myControl
)

# Print model to console
model
```

## Tuning model

Fit a random forest
As you saw in the video, random forest models are much more flexible than linear models, and can model complicated nonlinear effects as well as automatically capture interactions between variables. They tend to give very good results on real world data, so let's try one out on the wine quality dataset, where the goal is to predict the human-evaluated quality of a batch of wine, given some of the machine-measured chemical and physical properties of that batch.

Fitting a random forest model is exactly the same as fitting a generalized linear regression model, as you did in the previous chapter. You simply change the method argument in the train function to be "ranger". The ranger package is a rewrite of R's classic randomForest package and fits models much faster, but gives almost exactly the same results. We suggest that all beginners use the ranger package for random forest modeling.

```{r Frf}
wine <- read.csv("wine.csv")


# Fit random forest: model
model <- train(
  quality ~ .,
  tuneLength = 1,
  data = wine, 
  method = "ranger",
  trControl = trainControl(
    method = "cv", 
    number = 5, 
    verboseIter = TRUE
  )
)

# Print model to console
model
```

Try a longer tune length
Recall from the video that random forest models have a primary tuning parameter of mtry, which controls how many variables are exposed to the splitting search routine at each split. For example, suppose that a tree has a total of 10 splits and mtry = 2. This means that there are 10 samples of 2 predictors each time a split is evaluated.

Use a larger tuning grid this time, but stick to the defaults provided by the train() function. Try a tuneLength of 3, rather than 1, to explore some more potential models, and plot the resulting model using the plot function.

```{r Tltl}
# Fit random forest: model
model <- train(
  quality~.,
  tuneLength = 3,
  data = wine, 
  method = "ranger",
  trControl = trainControl(
    method = "cv", 
    number = 5, 
    verboseIter = TRUE
  )
)

# Print model to console
model

# Plot model
plot(model)
```

Fit a random forest with custom tuning
Now that you've explored the default tuning grids provided by the train() function, let's customize your models a bit more.

You can provide any number of values for mtry, from 2 up to the number of columns in the dataset. In practice, there are diminishing returns for much larger values of mtry, so you will use a custom tuning grid that explores 2 simple models (mtry = 2 and mtry = 3) as well as one more complicated model (mtry = 7).

```{r frfwct}
# Define the tuning grid: tuneGrid
tuneGrid <- data.frame(
  .mtry = c(2,3,7),
  .splitrule = "variance",
  .min.node.size = 5
)


# Fit random forest: model
model <- train(
  quality~.,
  tuneGrid = tuneGrid,
  data = wine, 
  method = "ranger",
  trControl = trainControl(
    method = "cv", 
    number = 5, 
    verboseIter = TRUE
  )
)

# Print model to console
model

# Plot model
plot(model)
```

Make a custom trainControl
The wine quality dataset was a regression problem, but now you are looking at a classification problem. This is a simulated dataset based on the "don't overfit" competition on Kaggle a number of years ago.

Classification problems are a little more complicated than regression problems because you have to provide a custom summaryFunction to the train() function to use the AUC metric to rank your models. Start by making a custom trainControl, as you did in the previous chapter. Be sure to set classProbs = TRUE, otherwise the twoClassSummary for summaryFunction will break.

```{r Mctc}
# Create custom trainControl: myControl
myControl <- trainControl(method = "cv", number = 10, summaryFunction = twoClassSummary, classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE)
```

Fit glmnet with custom trainControl
Now that you have a custom trainControl object, fit a glmnet model to the "don't overfit" dataset. Recall from the video that glmnet is an extension of the generalized linear regression model (or glm) that places constraints on the magnitude of the coefficients to prevent overfitting. This is more commonly known as "penalized" regression modeling and is a very useful technique on datasets with many predictors and few values.

glmnet is capable of fitting two different kinds of penalized models, controlled by the alpha parameter:

Ridge regression (or alpha = 0)
Lasso regression (or alpha = 1)
You'll now fit a glmnet model to the "don't overfit" dataset using the defaults provided by the caret package.

```{r fgctc}
overfit <- read.csv("overfit.csv")

# Fit glmnet model: model
model <- train(
  y ~ ., 
  overfit,
  method = "glmnet",
  trControl = myControl
)

# Print model to console
model

# Print maximum ROC statistic
max(model[["results"]][["ROC"]])
```

```{r gwctcat}
# Train glmnet with custom trainControl and tuning: model
model <- train(y~., data = overfit,
  tuneGrid = expand.grid(alpha = 0:1,lambda = seq(0.0001, 1, length = 20)),
  method = "glmnet",
  trControl = myControl
)

# Print model to console
model

# Print maximum ROC statistic
max(model[["results"]][["ROC"]])
```

## Preprocessing

Apply median imputation
In this chapter, you'll be using a version of the Wisconsin Breast Cancer dataset. This dataset presents a classic binary classification problem: 50% of the samples are benign, 50% are malignant, and the challenge is to identify which are which.

This dataset is interesting because many of the predictors contain missing values and most rows of the dataset have at least one missing value. This presents a modeling challenge, because most machine learning algorithms cannot handle missing values out of the box. For example, your first instinct might be to fit a logistic regression model to this data, but prior to doing this you need a strategy for handling the NAs.

Fortunately, the train() function in caret contains an argument called preProcess, which allows you to specify that median imputation should be used to fill in the missing values. In previous chapters, you created models with the train() function using formulas such as y ~ .. An alternative way is to specify the x and y arguments to train(), where x is an object with samples in rows and features in columns and y is a numeric or factor vector containing the outcomes. Said differently, x is a matrix or data frame that contains the whole dataset you'd use for the data argument to the lm() call, for example, but excludes the response variable column; y is a vector that contains just the response variable column.

For this exercise, the argument x to train() is loaded in your workspace as breast_cancer_x and y as breast_cancer_y.

```{r AMI}
breast_cancer_x <- read.csv("breast_cancer_x.csv") %>% select(-1)
breast_cancer_y <- read.csv("breast_cancer_y.csv") %>% select(-1)
breast_cancer_y <- as.vector(breast_cancer_y$x)


# Apply median imputation: median_model
median_model <- train(x = breast_cancer_x, y = breast_cancer_y, method = "glm", trControl = myControl,
  preProcess = "medianImpute")

# Print median_model to console
median_model
```

Use KNN imputation
In the previous exercise, you used median imputation to fill in missing values in the breast cancer dataset, but that is not the only possible method for dealing with missing data.

An alternative to median imputation is k-nearest neighbors, or KNN, imputation. This is a more advanced form of imputation where missing values are replaced with values from other rows that are similar to the current row. While this is a lot more complicated to implement in practice than simple median imputation, it is very easy to explore in caret using the preProcess argument to train(). You can simply use preProcess = "knnImpute" to change the method of imputation used prior to model fitting.

```{r KNNi}
# Apply KNN imputation: knn_model
knn_model <- train(x = breast_cancer_x, y = breast_cancer_y, method = "glm", trControl = myControl,
  preProcess = "knnImpute")

# Print knn_model to console
knn_model
```

Combining preprocessing methods
The preProcess argument to train() doesn't just limit you to imputing missing values. It also includes a wide variety of other preProcess techniques to make your life as a data scientist much easier. You can read a full list of them by typing ?preProcess and reading the help page for this function.

One set of preprocessing functions that is particularly useful for fitting regression models is standardization: centering and scaling. You first center by subtracting the mean of each column from each value in that column, then you scale by dividing by the standard deviation.

Standardization transforms your data such that for each column, the mean is 0 and the standard deviation is 1. This makes it easier for regression models to find a good solution.

```{r cpm}
# Fit glm with median imputation
model <- train(x = breast_cancer_x, y = breast_cancer_y, method = "glm", trControl = myControl,
  preProcess = "medianImpute")

# Print knn_model to console
model

# Update model with standardization
model <- train(x = breast_cancer_x, y = breast_cancer_y, method = "glm", trControl = myControl,
  preProcess = c("medianImpute","center","scale"))

# Print updated model
model
```

Remove near zero variance predictors
As you saw in the video, for the next set of exercises, you'll be using the blood-brain dataset. This is a biochemical dataset in which the task is to predict the following value for a set of biochemical compounds:

log((concentration of compound in brain) /
      (concentration of compound in blood))
This gives a quantitative metric of the compound's ability to cross the blood-brain barrier, and is useful for understanding the biological properties of that barrier.

One interesting aspect of this dataset is that it contains many variables and many of these variables have extemely low variances. This means that there is very little information in these variables because they mostly consist of a single value (e.g. zero).

Fortunately, caret contains a utility function called nearZeroVar() for removing such variables to save time during modeling.

nearZeroVar() takes in data x, then looks at the ratio of the most common value to the second most common value, freqCut, and the percentage of distinct values out of the number of total samples, uniqueCut. By default, caret uses freqCut = 19 and uniqueCut = 10, which is fairly conservative. I like to be a little more aggressive and use freqCut = 2 and uniqueCut = 20 when calling nearZeroVar().

```{r Rnzvp}
bloodbrain_x <- read.csv("bloodbrain_x.csv") %>% select(-1)
bloodbrain_y <- read.csv("bloodbrain_y.csv") %>% select(-1)
bloodbrain_y <- as.vector(bloodbrain_y$x)

# Identify near zero variance predictors: remove_cols
remove_cols <- nearZeroVar(bloodbrain_x, names = TRUE, 
                           freqCut = 2, uniqueCut = 20)

# Get all column names from bloodbrain_x: all_cols
all_cols <- names(bloodbrain_x)

# Remove from data: bloodbrain_x_small
bloodbrain_x_small <- bloodbrain_x[ , setdiff(all_cols, remove_cols)]
```

Fit model on reduced blood-brain data
Now that you've reduced your dataset, you can fit a glm model to it using the train() function. This model will run faster than using the full dataset and will yield very similar predictive accuracy.

Furthermore, zero variance variables can cause problems with cross-validation (e.g. if one fold ends up with only a single unique value for that variable), so removing them prior to modeling means you are less likely to get errors during the fitting process.

bloodbrain_x, bloodbrain_y, remove, and bloodbrain_x_small are loaded in your workspace.

Fit a glm model using the train() function and the reduced blood-brain dataset you created in the previous exercise.
Print the result to the console.

```{r fmrbbd}
remove <- remove_cols

# Fit model on reduced data: model
model <- train(x = bloodbrain_x_small, y = bloodbrain_y, method = "glm")

# Print model to console
model
```

Using PCA as an alternative to nearZeroVar()
An alternative to removing low-variance predictors is to run PCA on your dataset. This is sometimes preferable because it does not throw out all of your data: many different low variance predictors may end up combined into one high variance PCA variable, which might have a positive impact on your model's accuracy.

This is an especially good trick for linear models: the pca option in the preProcess argument will center and scale your data, combine low variance variables, and ensure that all of your predictors are orthogonal. This creates an ideal dataset for linear regression modeling, and can often improve the accuracy of your models.

```{r pcaanzv}
# Fit glm model using PCA: model
model <- train(x = bloodbrain_x, y = bloodbrain_y, method = "glm", preProcess = "pca")

# Print model to console
model
```

## Selecting Models

Make custom train/test indices
As you saw in the video, for this chapter you will focus on a real-world dataset that brings together all of the concepts discussed in the previous chapters.

The churn dataset contains data on a variety of telecom customers and the modeling challenge is to predict which customers will cancel their service (or churn).

In this chapter, you will be exploring two different types of predictive models: glmnet and rf, so the first order of business is to create a reusable trainControl object you can use to reliably compare them.

```{r Mctti}
churn_y <- read.csv("churn_y.csv")
churn_y <- as.vector(churn_y$x)

# Create custom indices: myFolds
myFolds <- createFolds(churn_y, k = 5)

# Create reusable trainControl object: myControl
myControl <- trainControl(summaryFunction = twoClassSummary, classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE, savePredictions = TRUE, index = myFolds)
```

Fit the baseline model
Now that you have a reusable trainControl object called myControl, you can start fitting different predictive models to your churn dataset and evaluate their predictive accuracy.

You'll start with one of my favorite models, glmnet, which penalizes linear and logistic regression models on the size and number of coefficients to help prevent overfitting.

```{r fbm}
churn_x <- read.csv("churn_x.csv") %>% select(-1)

# Fit glmnet model: model_glmnet
model_glmnet <- train(x = churn_x, y = churn_y, metric = "ROC", method = "glmnet", trControl = myControl)
```

Random forest with custom trainControl
Another one of my favorite models is the random forest, which combines an ensemble of non-linear decision trees into a highly flexible (and usually quite accurate) model.

Rather than using the classic randomForest package, you'll be using the ranger package, which is a re-implementation of randomForest that produces almost the exact same results, but is faster, more stable, and uses less memory. I highly recommend it as a starting point for random forest modeling in R.

```{r rfwctc}
# Fit random forest: model_rf
model_rf <- train(x = churn_x, y = churn_y, metric = "ROC", method = "ranger", trControl = myControl)
```

Create a resamples object
Now that you have fit two models to the churn dataset, it's time to compare their out-of-sample predictions and choose which one is the best model for your dataset.

You can compare models in caret using the resamples() function, provided they have the same training data and use the same trainControl object with preset cross-validation folds. resamples() takes as input a list of models and can be used to compare dozens of models at once (though in this case you are only comparing two models).

```{r cro}
# Create model_list
model_list <- list(item1 = model_glmnet, item2 = model_rf)

# Pass model_list to resamples(): resamples
resamples <- resamples(model_list)

# Summarize the results
summary(resamples)
```

Create a box-and-whisker plot
caret provides a variety of methods to use for comparing models. All of these methods are based on the resamples() function. My favorite is the box-and-whisker plot, which allows you to compare the distribution of predictive accuracy (in this case AUC) for the two models.

In general, you want the model with the higher median AUC, as well as a smaller range between min and max AUC.

You can make this plot using the bwplot() function, which makes a box and whisker plot of the model's out of sample scores. Box and whisker plots show the median of each distribution as a line and the interquartile range of each distribution as a box around the median line. You can pass the metric = "ROC" argument to the bwplot() function to show a plot of the model's out-of-sample ROC scores and choose the model with the highest median ROC.

If you do not specify a metric to plot, bwplot() will automatically plot 3 of them.

```{r cbwp}
# Create bwplot
bwplot(resamples, metric = "ROC")
```

Create a scatterplot
Another useful plot for comparing models is the scatterplot, also known as the xy-plot. This plot shows you how similar the two models' performances are on different folds.

It's particularly useful for identifying if one model is consistently better than the other across all folds, or if there are situations when the inferior model produces better predictions on a particular subset of the data.

```{r cs}
# Create xyplot
xyplot(resamples, metric = "ROC")
```

Ensembling models
That concludes the course! As a teaser for a future course on making ensembles of caret models, I'll show you how to fit a stacked ensemble of models using the caretEnsemble package.

caretEnsemble provides the caretList() function for creating multiple caret models at once on the same dataset, using the same resampling folds. You can also create your own lists of caret models.

In this exercise, I've made a caretList for you, containing the glmnet and ranger models you fit on the churn dataset. Use the caretStack() function to make a stack of caret models, with the two sub-models (glmnet and ranger) feeding into another (hopefully more accurate!) caret model.



