---
title: "Shinydashboard"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  pdf_document: default
  html_document:
    code_folding: hide
    fig_align: center
    fig_height: 6
    fig_width: 12
    keep_md: yes
editor_options:
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, shiny, shinydashboard)

theme_set(theme_bw())
```

https://www.datacamp.com/courses/building-dashboards-with-shinydashboard

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## Static Dashboards

Create empty Header, Sidebar, and Body
Using the functions discussed, create an "empty" header, sidebar, and body.

```{r 101}
library(shinydashboard)

# Create an empty header
header <- dashboardHeader()

# Create an empty sidebar
sidebar <- dashboardSidebar()

# Create an empty body
body <- dashboardBody()
```

Create an empty Shiny Dashboard
Using the header, sidebar, and body created in the previous exercise, create an "empty" dashboard with a header, sidebar, and body.

```{r 102}
header <- dashboardHeader()
sidebar <- dashboardSidebar()
body <- dashboardBody()

# Create the UI using the header, sidebar, and body
ui <- dashboardPage(header,sidebar,body)

server <- function(input, output) {}

shinyApp(ui, server)
```

Create message menus
Let's update your empty dashboard. You can add menus to your header using the dropdownMenu() function. One type of menu is the message menu, which can be created by adding type = "messages".

header <- dashboardHeader(
  dropdownMenu(type = "messages")
)
You are building a NASA themed app. So far, the application includes a message in the header drop down menu pointing viewers to where they can find out when the International Space Station can be seen overhead.

header <- dashboardHeader(
  dropdownMenu(
    type = "messages",
    messageItem(
        from = "Lucy",
        message = "You can view the International Space Station!",
        href = "https://spotthestation.nasa.gov/sightings/"
        )
    )
)
Add a second message using the messageItem() function to link viewers to the frequently asked questions: https://spotthestation.nasa.gov/faq.cfm

We've already defined an empty server function for you.

```{r 103}
header <- dashboardHeader(
    dropdownMenu(
    type = "messages",
    messageItem(
        from = "Lucy",
        message = "You can view the International Space Station!",
        href = "https://spotthestation.nasa.gov/sightings/"
        ),
    # Add a second messageItem() 
    messageItem(
        from = "Lucy",
        message = "Learn more about the International Space Station",
        href = "https://spotthestation.nasa.gov/faq.cfm"
        )
    )
)

ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create notification menus
You can also create a notification drop down menu on your header. This allows you to notify your user of something. Add a notification that the International Space Station is overhead.

```{r 104}
header <- dashboardHeader(
  # Create a notification drop down menu
  dropdownMenu(
    type = "notifications",
    notificationItem(
      text = "The International Space Station is overhead!"
    )
  )
)

# Use the new header
ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create task menus
A third type of drop down menu is the tasks menu. In addition to providing text, these items show progress on a task using the value parameter. For example, if you wanted to indicate that the task "Alert users to the International Space Station" was 80% complete, you would indicate it in a taskItem() in the following manner.

taskItem(
  text = "Alert users to the International Space Station",
  value = 80
    )
Add a tasks drop down menu to the header that indicates the task "Mission Learn Shiny Dashboard" is 10% complete.

We've already defined an empty server function for you.

```{r 105}
header <- dashboardHeader(
  dropdownMenu(
    # Create a tasks drop down menu
    type = "tasks",
    taskItem(
      text = "Mission Learn Shiny Dashboard",
      value = 10
    )
  )
)

# Use the new header
ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody())
shinyApp(ui, server)
```

Create Sidebar tabs
You can create tabs for you dashboard. First, create tabs on the sidebar using the sidebarMenu() function. You will create two tabs, one for the "Dashboard" and one for the "Inputs".

```{r 106}
sidebar <- dashboardSidebar(
  sidebarMenu(
  # Create two `menuItem()`s, "Dashboard" and "Inputs"
    menuItem(text = "Dashboard",
            tabName = "dashboard"
            ), 
    menuItem(text = "Inputs", 
            tabName = "inputs"
            )
  )
)

# Use the new sidebar
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create Body tabs
Now that you have created tabs on the sidebar, link them to tabs in the body using the tabItems() function in the dashboardBody().

```{r 107}
body <- dashboardBody(
  tabItems(
  # Add two tab items, one with `tabName` "dashboard" and one with `tabName` "inputs"
    tabItem(tabName = "dashboard"),
    tabItem(tabName = "inputs")
  )
)

# Use the new body
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

Create tab boxes
Now that you have created tabs on the sidebar, and tabs in the body, we can add boxes within each of the body's tabs (and there can be tabs within the boxes!) You can add a tabBox() directly to the dashboardBody() or place it within a tabItem(). Here is an example of a tabBox() directly in the dashboardBody(). The tabPanel() function is from the shiny package and will add tabs within the box.

body <-  dashboardBody(
    tabBox(
      title = "Tracking the International Space Station",
      tabPanel("Tab1", "Content for the first tab"),
      tabPanel("Tab2", "Content for the second tab")
    )
  )
Add a tabBox() to the dashboard body you have been working on in the previous exercises.

```{r 108}
library("shiny")
body <- dashboardBody(
  # Create a tabBox
  tabItems(
    tabItem(
      tabName = "dashboard",
      tabBox(
        title = "International Space Station Fun Facts",
        tabPanel("Fun Fact 1"),
        tabPanel("Fun Fact 2")
        )
    ),
    tabItem(tabName = "inputs")
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

Review selectInput and sliderInput
As a quick refresher, using the dashboardSidebar() function along with selectInput() and sliderInput() from the shiny package, you can create select lists and sliders. For example, to create a select list in the sidebar, you can do the following:

sidebar <- dashboardSidebar(
  selectInput(inputId = "numbers",
              label = "Numbers",
              choices = 1:3)
  )
For this chapter, we are working with the dplyr starwars dataset, a data set pulled from the Star Wars API containing information about characters from Star Wars. We are interested in creating a subset of the data set of characters between specified height parameters. For this exercise, create a slider to indicate the maximum height you'd like in your subset.

```{r 201}
sidebar <- dashboardSidebar(
  # Add a slider
  sliderInput(
    inputId = "height",
    label = "Height",
    min = 66,
    max = 264,
    value = 264)
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = dashboardBody()
                    )
shinyApp(ui, server) 
```

Reactive expression practice
Using the dplyr starwars dataset, create a shiny app that allows the user to select a character by name and output the selected name in the body. The data frame is already loaded as starwars.

```{r 202}
library(shiny)
sidebar <- dashboardSidebar(
  # Create a select list
  selectInput(inputId = "name", 
              label = "Name",
              choices = starwars$name
  )
)
body <- dashboardBody(
  textOutput("name")
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )

server <- function(input, output) {
  output$name <- renderText({
      input$name
    })
}

shinyApp(ui, server)
```

Read in real-time data
One benefit of a dashboard is the ability to examine real-time data. In shiny, this can be done using the reactiveFileReader() or reactivePoll() functions from the shiny package. For example, if we had our data saved as a .csv, we could read it in using reactiveFileReader() along with the readFunc set to read.csv.

filepath <- "data.csv"

server <- function(input, output, session) {
  reactive_data <- reactiveFileReader(
    intervalMillis = 1000,
    session = session, 
    filePath = filepath,
    readFunc = read.csv
  )
}
We have our data saved as a .csv located at a url called starwars_url; this object is already loaded for your convenience. In order to read this in, we can set our own readFunc like this:

readFunc = function(filePath) { 
  read.csv(url(filePath))
  }
If this .csv were updated live, we would see the changes! Read in this real-time data using reactiveFileReader().

```{r 203}
library("shiny")

starwars_url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_6225/datasets/starwars.csv"

server <- function(input, output, session) {
  reactive_starwars_data <- reactiveFileReader(
         intervalMillis = 1000,
         session =  session,
         filePath =  starwars_url,
         readFunc = function(filePath) { 
           read.csv(url(filePath))
         }
        )
    }
  
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

View real-time data
Now that you've read in the real-time data as reactive_starwars_data(), you can examine it as it updates using renderTable(). If you save this reactive table, you can then render it using the tableOutput() function. For example, if we had a reactive data frame called my_reactive_data() we could save this as output$my_table and render it using tableOutput() with the following code:

server <- function(input, output, session) {
  output$my_table <- renderTable({
    my_reactive_data()
  })
}

body <- dashboardBody(
  tableOutput("my_table")
)

```{r 204}
library(shiny)

server <- function(input, output, session) {
  reactive_starwars_data <- reactiveFileReader(
        intervalMillis = 1000,
        session = session,
        filePath = starwars_url,
        readFunc = function(filePath) { 
           read.csv(url(filePath))
         }
         )
  
  output$table <- renderTable({
    reactive_starwars_data()
  })
}

body <- dashboardBody(
  tableOutput("table")
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```






















