---
title: "Shinydashboard"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  pdf_document: default
  html_document:
    code_folding: hide
    fig_align: center
    fig_height: 6
    fig_width: 12
    keep_md: yes
editor_options:
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, shiny, shinydashboard)

theme_set(theme_bw())
```

https://www.datacamp.com/courses/building-dashboards-with-shinydashboard

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## Static Dashboards

Create empty Header, Sidebar, and Body
Using the functions discussed, create an "empty" header, sidebar, and body.

```{r 101}
library(shinydashboard)

# Create an empty header
header <- dashboardHeader()

# Create an empty sidebar
sidebar <- dashboardSidebar()

# Create an empty body
body <- dashboardBody()
```

Create an empty Shiny Dashboard
Using the header, sidebar, and body created in the previous exercise, create an "empty" dashboard with a header, sidebar, and body.

```{r 102}
header <- dashboardHeader()
sidebar <- dashboardSidebar()
body <- dashboardBody()

# Create the UI using the header, sidebar, and body
ui <- dashboardPage(header,sidebar,body)

server <- function(input, output) {}

shinyApp(ui, server)
```

Create message menus
Let's update your empty dashboard. You can add menus to your header using the dropdownMenu() function. One type of menu is the message menu, which can be created by adding type = "messages".

header <- dashboardHeader(
  dropdownMenu(type = "messages")
)
You are building a NASA themed app. So far, the application includes a message in the header drop down menu pointing viewers to where they can find out when the International Space Station can be seen overhead.

header <- dashboardHeader(
  dropdownMenu(
    type = "messages",
    messageItem(
        from = "Lucy",
        message = "You can view the International Space Station!",
        href = "https://spotthestation.nasa.gov/sightings/"
        )
    )
)
Add a second message using the messageItem() function to link viewers to the frequently asked questions: https://spotthestation.nasa.gov/faq.cfm

We've already defined an empty server function for you.

```{r 103}
header <- dashboardHeader(
    dropdownMenu(
    type = "messages",
    messageItem(
        from = "Lucy",
        message = "You can view the International Space Station!",
        href = "https://spotthestation.nasa.gov/sightings/"
        ),
    # Add a second messageItem() 
    messageItem(
        from = "Lucy",
        message = "Learn more about the International Space Station",
        href = "https://spotthestation.nasa.gov/faq.cfm"
        )
    )
)

ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create notification menus
You can also create a notification drop down menu on your header. This allows you to notify your user of something. Add a notification that the International Space Station is overhead.

```{r 104}
header <- dashboardHeader(
  # Create a notification drop down menu
  dropdownMenu(
    type = "notifications",
    notificationItem(
      text = "The International Space Station is overhead!"
    )
  )
)

# Use the new header
ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create task menus
A third type of drop down menu is the tasks menu. In addition to providing text, these items show progress on a task using the value parameter. For example, if you wanted to indicate that the task "Alert users to the International Space Station" was 80% complete, you would indicate it in a taskItem() in the following manner.

taskItem(
  text = "Alert users to the International Space Station",
  value = 80
    )
Add a tasks drop down menu to the header that indicates the task "Mission Learn Shiny Dashboard" is 10% complete.

We've already defined an empty server function for you.

```{r 105}
header <- dashboardHeader(
  dropdownMenu(
    # Create a tasks drop down menu
    type = "tasks",
    taskItem(
      text = "Mission Learn Shiny Dashboard",
      value = 10
    )
  )
)

# Use the new header
ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody())
shinyApp(ui, server)
```

Create Sidebar tabs
You can create tabs for you dashboard. First, create tabs on the sidebar using the sidebarMenu() function. You will create two tabs, one for the "Dashboard" and one for the "Inputs".

```{r 106}
sidebar <- dashboardSidebar(
  sidebarMenu(
  # Create two `menuItem()`s, "Dashboard" and "Inputs"
    menuItem(text = "Dashboard",
            tabName = "dashboard"
            ), 
    menuItem(text = "Inputs", 
            tabName = "inputs"
            )
  )
)

# Use the new sidebar
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create Body tabs
Now that you have created tabs on the sidebar, link them to tabs in the body using the tabItems() function in the dashboardBody().

```{r 107}
body <- dashboardBody(
  tabItems(
  # Add two tab items, one with `tabName` "dashboard" and one with `tabName` "inputs"
    tabItem(tabName = "dashboard"),
    tabItem(tabName = "inputs")
  )
)

# Use the new body
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

Create tab boxes
Now that you have created tabs on the sidebar, and tabs in the body, we can add boxes within each of the body's tabs (and there can be tabs within the boxes!) You can add a tabBox() directly to the dashboardBody() or place it within a tabItem(). Here is an example of a tabBox() directly in the dashboardBody(). The tabPanel() function is from the shiny package and will add tabs within the box.

body <-  dashboardBody(
    tabBox(
      title = "Tracking the International Space Station",
      tabPanel("Tab1", "Content for the first tab"),
      tabPanel("Tab2", "Content for the second tab")
    )
  )
Add a tabBox() to the dashboard body you have been working on in the previous exercises.

```{r 108}
library("shiny")
body <- dashboardBody(
  # Create a tabBox
  tabItems(
    tabItem(
      tabName = "dashboard",
      tabBox(
        title = "International Space Station Fun Facts",
        tabPanel("Fun Fact 1"),
        tabPanel("Fun Fact 2")
        )
    ),
    tabItem(tabName = "inputs")
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

## Dynamic

Review selectInput and sliderInput
As a quick refresher, using the dashboardSidebar() function along with selectInput() and sliderInput() from the shiny package, you can create select lists and sliders. For example, to create a select list in the sidebar, you can do the following:

sidebar <- dashboardSidebar(
  selectInput(inputId = "numbers",
              label = "Numbers",
              choices = 1:3)
  )
For this chapter, we are working with the dplyr starwars dataset, a data set pulled from the Star Wars API containing information about characters from Star Wars. We are interested in creating a subset of the data set of characters between specified height parameters. For this exercise, create a slider to indicate the maximum height you'd like in your subset.

```{r 201}
sidebar <- dashboardSidebar(
  # Add a slider
  sliderInput(
    inputId = "height",
    label = "Height",
    min = 66,
    max = 264,
    value = 264)
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = dashboardBody()
                    )
shinyApp(ui, server) 
```

Reactive expression practice
Using the dplyr starwars dataset, create a shiny app that allows the user to select a character by name and output the selected name in the body. The data frame is already loaded as starwars.

```{r 202}
library(shiny)
sidebar <- dashboardSidebar(
  # Create a select list
  selectInput(inputId = "name", 
              label = "Name",
              choices = starwars$name
  )
)
body <- dashboardBody(
  textOutput("name")
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )

server <- function(input, output) {
  output$name <- renderText({
      input$name
    })
}

shinyApp(ui, server)
```

Read in real-time data
One benefit of a dashboard is the ability to examine real-time data. In shiny, this can be done using the reactiveFileReader() or reactivePoll() functions from the shiny package. For example, if we had our data saved as a .csv, we could read it in using reactiveFileReader() along with the readFunc set to read.csv.

filepath <- "data.csv"

server <- function(input, output, session) {
  reactive_data <- reactiveFileReader(
    intervalMillis = 1000,
    session = session, 
    filePath = filepath,
    readFunc = read.csv
  )
}
We have our data saved as a .csv located at a url called starwars_url; this object is already loaded for your convenience. In order to read this in, we can set our own readFunc like this:

readFunc = function(filePath) { 
  read.csv(url(filePath))
  }
If this .csv were updated live, we would see the changes! Read in this real-time data using reactiveFileReader().

```{r 203}
library("shiny")

starwars_url <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_6225/datasets/starwars.csv"

server <- function(input, output, session) {
  reactive_starwars_data <- reactiveFileReader(
         intervalMillis = 1000,
         session =  session,
         filePath =  starwars_url,
         readFunc = function(filePath) { 
           read.csv(url(filePath))
         }
        )
    }
  
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

View real-time data
Now that you've read in the real-time data as reactive_starwars_data(), you can examine it as it updates using renderTable(). If you save this reactive table, you can then render it using the tableOutput() function. For example, if we had a reactive data frame called my_reactive_data() we could save this as output$my_table and render it using tableOutput() with the following code:

server <- function(input, output, session) {
  output$my_table <- renderTable({
    my_reactive_data()
  })
}

body <- dashboardBody(
  tableOutput("my_table")
)

```{r 204}
library(shiny)

server <- function(input, output, session) {
  reactive_starwars_data <- reactiveFileReader(
        intervalMillis = 1000,
        session = session,
        filePath = starwars_url,
        readFunc = function(filePath) { 
           read.csv(url(filePath))
         }
         )
  
  output$table <- renderTable({
    reactive_starwars_data()
  })
}

body <- dashboardBody(
  tableOutput("table")
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Create reactive menu items
We have added dynamic content via subsetting a data frame based on using input and reading in real-time data. Now we are going to allow the user to input task_data to determine task items. Recall that we can use an apply() function to iterate over a data frame, applying the taskItem() function to each row.

tasks <- apply(task_data, 1, function(row) { 
  taskItem(text = row[["text"]],
           value = row[["value"]])
})
dropdownMenu(type = "tasks", .list = tasks)
You have a data frame (already loaded) called task_data with columns text and value. Use this to create a task drop down menu.

```{r 205}
task_data <- read.csv("task_data.csv")

server <- function(input, output) {
  output$task_menu <- renderMenu({
      tasks <- apply(task_data, 1, function(row) {
        taskItem(text = row[["text"]],
                 value = row[["value"]])
                 })
    
      dropdownMenu(type = "tasks", .list = tasks)
  })
}

header <- dashboardHeader(dropdownMenuOutput("task_menu"))

ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Create reactive boxes
In addition to updating the task drop down menus reactively, it is possible to create reactive boxes in the body. Here, you have a clickable action button in your sidebar. You want to update a value box each time the user clicks the action button. The valueBox() function will create a value box in the body. This takes the following form:

valueBox(value = 10,
        subtitle = "There are ten things here!"
        )
        
```{r 206}
library("shiny")

sidebar <- dashboardSidebar(
  actionButton("click", "Update click box")
) 

server <- function(input, output) {
  output$click_box <- renderValueBox({
    valueBox(
      input$click, 
     "Click Box"
    )
  })
}

body <- dashboardBody(
      valueBoxOutput("click_box")
 )

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

## Customize

Create body with row-based layout
Recall that you can create a dashboard with row-based layout using the fluidRow() function from shiny within the dashboardBody(). We want to create two rows, one with the following content:

# Row 1
  box(
    width = 12,
    title = "Regular Box, Row 1",
    "Star Wars"
    )
and a second row with the following content:

# Row 2
  box(
    width = 12,
    title = "Regular Box, Row 2",
    "Nothing but Star Wars"
    )
    
```{r 301}
library("shiny")

body <- dashboardBody(
  fluidRow(
# Row 1
    box(
      width = 12,
      title = "Regular Box, Row 1",
      "Star Wars"
    )
  ),
  fluidRow(
# Row 2
    box(
      width = 12,
      title = "Regular Box, Row 2",
      "Nothing but Star Wars"
    )
  )
)


ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```
 
Create body with column-based layout
Recall that you can create your contents in columns using a combination of the fluidRow() and column() functions from the shiny package. For example, for two columns of content in the body, you can do the following:

body <- dashboardBody(
  fluidRow(
    column(width = 6,
      "This is one column"
      ),
    column(width = 6,
      "This is a second column"
    )
  )
)
Let's create two columns where the first column includes the following content:

# Column 1
  infoBox(
    width = NULL,
    title = "Regular Box, Column 1",
    subtitle = "Gimme those Star Wars"
    )
and a second column with the following content:

# Column 2
  infoBox(
    width = NULL,
    title = "Regular Box, Column 2",
    subtitle = "Don't let them end"
    )
Recall that with column layouts, you want to set the width = NULL within the boxes (this is true for box(), infoBox(), and valueBox()), and control the width with the column() function.

```{r 302}
library("shiny")

body <- dashboardBody(
  fluidRow(
# Column 1
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Column 1",
        subtitle = "Gimme those Star Wars"
      )
    ),
# Column 2
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Column 2",
        subtitle = "Don't let them end"
      )
    )
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Create body with mixed layout
In addition to having layouts specified using rows and columns, we can have mixed layouts that combine the two.

Let's create two rows where the first row includes the following content:

# Row 1
  box(
    width = 12,
    title = "Regular Box, Row 1",
    "Star Wars, nothing but Star Wars"
    )
and a second row has two columns, where each column has the following content:

# Row 2, Column 1
  infoBox(
    width = NULL,
    title = "Regular Box, Row 2, Column 1",
    subtitle = "Gimme those Star Wars"
    )
# Row 2, Column 2
  infoBox(
    width = NULL,
    title = "Regular Box, Row 2, Column 2",
    subtitle = "Don't let them end"
    )
    
```{r 303}
library("shiny")

body <- dashboardBody(
  fluidRow(
# Row 1
    box(
      width = 12,
      title = "Regular Box, Row 1",
      "Star Wars, nothing but Star Wars"
    )
  ),
  fluidRow(
# Row 2, Column 1
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 1",
        subtitle = "Gimme those Star Wars"
    )
   ),
# Row 2, Column 2
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 2",
        subtitle = "Don't let them end"
    )
  )
 )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Change the appearance of the dashboard
Walt Disney's favorite color is purple (mine too!). Using the application you've just created, let's change the color by updating the skin parameter to "purple". The body you just created is already loaded.

```{r 304}
# Update the skin
ui <- dashboardPage(
  skin = "purple",
  header = dashboardHeader(),
  sidebar = dashboardSidebar(),
  body = body)
  
# Run the app
shinyApp(ui, server)
```

Customize the body with CSS
You can customize the CSS by using the tags() function, for example you can add css to headers in the body in the following manner.

body <- dashboardBody(
    tags$head(
        tags$style(
            HTML('
            \\Add CSS here
            ')
        )
    )
)
Add the following CSS to make the box title bold by updating the font in the h3 header in the body:

h3 {
    font-weight: bold;
}

```{r 305}
library("shiny")

body <- dashboardBody(
# Update the CSS
  tags$head(
    tags$style(
      HTML('
      h3 {
        font-weight: bold;
      }
      ')
    )
  ),
  fluidRow(
    box(
      width = 12,
      title = "Regular Box, Row 1",
      "Star Wars, nothing but Star Wars"
    )
  ),
  fluidRow(
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 1",
        subtitle = "Gimme those Star Wars"
    )
   ),
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 2",
        subtitle = "Don't let them end"
    )
  )
 )
)

ui <- dashboardPage(
  skin = "purple",
  header = dashboardHeader(),
  sidebar = dashboardSidebar(),
  body = body)
shinyApp(ui, server)
```

Incorporate icons
Icons are a great way to make your dashboard look more polished! Remember the notification menu we created in Chapter 1? Spruce it up with a rocket icon! You can find other possible icons on the Font Awesome website. Recall that icons can be rendered using the icon() function in shiny; for example, the following will render a coffee icon:

icon("coffee")
The Font Awesome name for the rocket icon is "rocket".

```{r 306}
library("shiny")

header <- dashboardHeader(
  dropdownMenu(
    type = "notifications",
    notificationItem(
      text = "The International Space Station is overhead!",
      icon = icon("rocket")
    )
  )
)
ui <- dashboardPage(header = header,
                    sidebar = dashboardSidebar(),
                    body = dashboardBody()
                    )
shinyApp(ui, server)
```

Add some life to your layouts
In this exercise you'll jazz up a shinydashboard by adding some color. There's a dashboard with three different boxes already written for you. All you have to do is make the indicated changes to the boxes.

It might be useful to know that some valid statuses and their corresponding colors are:

primary Blue (sometimes dark blue)
success Green
info Blue
warning Orange
danger Red

```{r 307}
library("shiny")

body <- dashboardBody(
  tags$head(
    tags$style(
      HTML('
      h3 {
        font-weight: bold;
      }
      ')
    )
  ),
  fluidRow(
    box(
      width = 12,
      title = "Regular Box, Row 1",
      "Star Wars, nothing but Star Wars",
# Make the box red
      status = "danger"
    )
  ),
  fluidRow(
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 1",
        subtitle = "Gimme those Star Wars",
# Change the icon to a star
        icon = icon("star")
    )
   ),
    column(width = 6,
      infoBox(
        width = NULL,
        title = "Regular Box, Row 2, Column 2",
        subtitle = "Don't let them end",
# Make the box yellow
        color = "yellow"
    )
  )
 )
)

ui <- dashboardPage(
  skin = "purple",
  header = dashboardHeader(),
  sidebar = dashboardSidebar(),
  body = body)
shinyApp(ui, server)
```

## Case Study

Examine the variables in the data set
For the following exercises we are using the nasa_fireball data, a data set of reported fireball events. Explore the variables in this data set. The nasa_fireball data frame is already loaded.

```{r 401}
# Print the nasa_fireball data frame
nasa_fireball <- read.csv("nasa_fireball.csv")

# Examine the types of variables present
sapply(nasa_fireball, class)

# Observe the number of observations in this data frame
nrow(nasa_fireball)

# Check for missing data
sapply(nasa_fireball, anyNA)
```

Create a value box for the maximum velocity
Now you can begin building your dashboard! First, create a value box for the maximum velocity of a fireball at peak brightness (the variable vel in your nasa_fireball data frame). There are many ways to find the maximum value in R; be sure to remove NA values when doing so.

The nasa_fireball data frame has already been loaded.

```{r 402}
library("shiny")
max_vel <- max(nasa_fireball$vel, na.rm = TRUE)

body <- dashboardBody(
  fluidRow(
    # Add a value box for maximum velocity
    valueBox(
      value = max_vel,
      subtitle = "Maximum pre-impact velocity", 
      icon = icon("fire")
    )
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Create a value box for the maximum impact
Let's add another value box for the maximum log total impact energy. The variable in the nasa_fireball data frame corresponding to this is impact_e. The nasa_fireball data frame is again already loaded.

```{r 403}
library("shiny")

max_impact_e <- max(nasa_fireball$impact_e)

body <- dashboardBody(
  fluidRow(
    # Add a value box for maximum impact
    valueBox(
      value = max_impact_e, 
      subtitle = "Maximum impact energy (kilotons of TNT)",
      icon = icon("star")
    ),
    valueBox(
      value = max_vel,
      subtitle = "Maximum pre-impact velocity", 
      icon = icon("fire")
    )
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Create a value box for the maximum energy
Let's create a third value box for the maximum total optical radiated energy. The variable for this in the nasa_fireball data frame is energy. Again, the nasa_fireball data frame is loaded.

```{r 404}
library("shiny")

max_energy <- max(nasa_fireball$energy)

body <- dashboardBody(
  fluidRow(
    # Add a value box for maximum energy
    valueBox(
      value = max_energy, 
      subtitle = "Maximum total radiated energy (Joules)", 
      icon = icon("lightbulb-o")
    ),
    valueBox(
      value = max_impact_e, 
      subtitle = "Maximum impact energy (kilotons of TNT)",
      icon = icon("star")
    ),
    valueBox(
      value = max_vel,
      subtitle = "Maximum pre-impact velocity", 
      icon = icon("fire")
    )
  )
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Make the value boxes dynamic
The data frame contains the latitude and longitude of the location of the fireballs. Begin to build a dynamic value box containing the number of fireballs in the US. The approximate range of latitude for US cities is 19.5 to 64.9 and the range for longitude is -161.8 to -68.0. Count the number of fireballs in the current data frame that fall within this range. This count can be obtained a variety of ways in R. You may use the code below, or write you own and check it with the one provided.

n_us <- sum(
  ifelse(
    nasa_fireball$lat < 64.9 & nasa_fireball$lat > 19.5
      & nasa_fireball$lon < -68.0 & nasa_fireball$lon > -161.8,
    1, 0),
  na.rm = TRUE)
Then use the color parameter such that if n_us is less than 10, the value box is blue, otherwise it is fuchsia. The following logic can be used to determine the color.

if (n_us < 10) {
  "blue"
} else {
  "fuchsia"
}

```{r 405}
n_us <- sum(
  ifelse(
    nasa_fireball$lat < 64.9 & nasa_fireball$lat > 19.5
      & nasa_fireball$lon < -68.0 & nasa_fireball$lon > -161.8,
    1, 0),
  na.rm = TRUE)
  
server <- function(input, output) {
  output$us_box <- renderValueBox({
    valueBox(
      value = n_us,
      subtitle = "Number of Fireballs in the US",
      icon = icon("globe"),
      color = if (n_us < 10) {
                "blue"
              } else {
              "fuchsia"
              }
    )
  })
}

body <- dashboardBody(
  fluidRow(
    valueBoxOutput("us_box")
  )
)
ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = dashboardSidebar(),
                    body = body
                    )
shinyApp(ui, server)
```

Allow the user to input an alert threshold
Perhaps the people using your application would rather have a different threshold for changing the color of your value box. Create a slider in the sidebar to allow the user to input the threshold for the value box color.

```{r 406}
sidebar <- dashboardSidebar(
  sliderInput(
    inputId = "threshold",
    label = "Color Threshold",
    min = 0,
    max = 100,
    value = 10)
)

server <- function(input, output) {
  output$us_box <- renderValueBox({
    valueBox(
      value = n_us,
      subtitle = "Number of Fireballs in the US",
      icon = icon("globe"),
      color = if (n_us < input$threshold) {
                "blue"
              } else {
              "fuchsia"
              }
    )
  })
}


ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )
shinyApp(ui, server)
```

Create a dynamic plot of the location of fireballs
Rather than just examining how many fireballs have landed in the United States, you can plot each of the fireballs on a map! Here is some example code, for an example data set named df we can plot the fireball incidents as circles with the radius of the circle based on the velocity.

 output$plot <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%  
      addCircleMarkers(lng = df$lon, lat = df$lat, radius = df$vel, label = df$date, weight = 3)
    })
This code belongs in the server function. Using code similar to that above with the nasa_fireball data frame, create a plot of the fireball events with the radius equal to the log of the impact energy. To remove any warnings, I have subset the nasa_fireball data frame to remove any rows missing latitude or longitude.

```{r 407}
library("leaflet")

server <- function(input, output) {
  output$plot <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%  
      addCircleMarkers(
        lng = nasa_fireball$lon,
        lat = nasa_fireball$lat, 
        radius = log(nasa_fireball$impact_e), 
        label = nasa_fireball$date, 
        weight = 2)
    })
}

body <- dashboardBody( 
 leafletOutput("plot")
)

ui <- dashboardPage(
  header = dashboardHeader(),
  sidebar = dashboardSidebar(),
  body = body
)

shinyApp(ui, server)
```

Update the look of your new application
Now let's put it all together. You have created an application for viewing fireball locations. The full application is shown in the sample code. Update the skin of the dashboard page to a color of your choosing.

```{r 408}
library("leaflet")

server <- function(input, output) {
  output$plot <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%  
      addCircleMarkers(
        lng = nasa_fireball$lon,
        lat = nasa_fireball$lat, 
        radius = log(nasa_fireball$impact_e), 
        label = nasa_fireball$date, 
        weight = 2)
    })
}

body <- dashboardBody(
 fluidRow(
    valueBox(
      value = max_energy, 
      subtitle = "Maximum total radiated energy (Joules)", 
      icon = icon("lightbulb-o")
    ),
    valueBox(
      value = max_impact_e, 
      subtitle = "Maximum impact energy (kilotons of TNT)",
      icon = icon("star")
    ),
    valueBox(
      value = max_vel,
      subtitle = "Maximum pre-impact velocity", 
      icon = icon("fire")
    )
  ),
  fluidRow(
    leafletOutput("plot")
  )
)

ui <- dashboardPage(
  skin = "purple",
  header = dashboardHeader(),
  sidebar = dashboardSidebar(),
  body = body
)

shinyApp(ui, server)
```



