---
title: "Machine Learning Tidyverse"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

pacman::p_load(tidyverse)

```

```{r eval = FALSE, echo = FALSE}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
```


## Background

https://campus.datacamp.com/courses/communicating-with-data-in-the-tidyverse/custom-ggplot2-themes?ex=2

## ggplot2

In this chapter, you will have a first look at the data you're going to work with throughout this course: the relationship between weekly working hours and monetary compensation in European countries, according to the International Labour Organization (ILO). After that, you'll dive right in and discover a stunning correlation by employing an exploratory visualization. You will then apply a custom look to that graphic – you'll turn an ordinary plot into an aesthetically pleasing and unique data visualization.

Join the two data sets together
In the video, you have learned that the inner_join() function of dplyr needs to be given a "key" on which two data frames are joined. Actually, multiple keys that need to match can be specified. In this first exercise, you are going to join two data sets by two keys. The data frames ilo_hourly_compensation and ilo_working_hours are already loaded for you and are available in your workspace.

```{r jtdst}
ilo_hourly_compensation <- read.csv("ilo_hourly_compensation.csv") %>% select(-1)
ilo_working_hours <- read.csv("ilo_working_hours.csv") %>% select(-1)

# Join both data frames
ilo_data <- ilo_hourly_compensation %>% inner_join(ilo_working_hours, by = c("country", "year"))

# Count the resulting rows
ilo_data  %>% 
    count()

# Examine ilo_data
ilo_data
```

Change variable types
For displaying data in ggplot2 graphics, it is often helpful to convert all the variables to the right data type. Usually, categorical variables like country in this example should be converted to factors before plotting them. You can do so using as.factor(). In your data set, two columns are still of type "character" – use mutate() to turn them into factors.

```{r cvt}
# Turn year and country into a factor
ilo_data_corrected <- ilo_data %>%
  mutate(year = as.factor(as.numeric(year)), country = as.factor(country))
```

Filter the data for plotting
Use the filter() function of dplyr to remove non-European countries. A vector called european_countries has been created that contains all European countries which should remain in the data set. Here, the %in% operator introduced in the video might come in handy. Note that only a subset of European countries for which enough data exist is retained. For instance, Poland is missing.

```{r fdp}
european_countries <- readLines("european_countries.csv") 

# Examine the European countries vector
european_countries

# Only retain European countries
ilo_data <- ilo_data %>% filter(country %in% european_countries)
```

Some summary statistics
Use dplyrs group_by() and summarize() to compute summary statistics for both years.

```{r sss}
# Examine the structure of ilo_data
str(ilo_data)

# Group and summarize the data
ilo_data %>%
  group_by(year) %>%
  summarize(mean_hourly_compensation = mean(hourly_compensation),
            mean_working_hours = mean(working_hours))
```

A basic scatter plot
In this exercise, you will create a very basic scatter plot with ggplot2. This is mostly a repetition of stuff you've learnt in the prerequisites for this course, so it should be easy for you.

```{r bsp}
# Filter for 2006
plot_data <- ilo_data %>% filter(year == "2006")
  
# Create the scatter plot
ggplot(plot_data) +
  geom_point(aes(x = working_hours, y = hourly_compensation))
```

Add labels to the plot
As mentioned in the video, you're going to enhance the plot from the previous exercise by adding a title, a subtitle, and a caption to the plot as well as giving the axes meaningful names. You're going to use the labs() function to do this – try using ?labs in the console to see the available options.

```{r alp}
# Create the plot
ggplot(plot_data) +
  geom_point(aes(x = working_hours, y = hourly_compensation)) +
  # Add labels
  labs(
    x = "Working hours per week",
    y = "Hourly compensation",
    title = "The more people work, the less compensation they seem to receive",
    subtitle = "Working hours and hourly compensation in European countries, 2006",
    caption = "Data source: ILO, 2017"
  )
```





















