---
title: "Machine Learning Trees"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, DT, rpart, DMwR, caret, rpart.plot, rattle)
theme_set(theme_bw())
```

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

## Classification Trees

Build a classification tree
Let's get started and build our first classification tree. A classification tree is a decision tree that performs a classification (vs regression) task.

You will train a decision tree model to understand which loan applications are at higher risk of default using a subset of the German Credit Dataset. The response variable, called "default", indicates whether the loan went into a default or not, which means this is a binary classification problem (there are just two classes).

You will use the rpart package to fit the decision tree and the rpart.plot package to visualize the tree.

Instructions
100 XP
The data frame creditsub is in the workspace. This data frame is a subset of the original German Credit Dataset, which we will use to train our first classification tree model.

Take a look at the data using the str() function.
In R, formulas are used to model the response as a function of some set of predictors, so the formula here is default ~ ., which means use all columns (except the response column) as predictors.
Fit the classification decision tree using the rpart() function from the rpart package. In the rpart() function, note that you'll also have to provide the training data frame.
Using the model object that you create, plot the decision tree model using the rpart.plot() function from the rpart.plot package.


```{r bct}
creditsub <- read.csv("creditsub.csv")

# Look at the data
str(creditsub)

# Create the model
credit_model <- rpart(formula = default ~ ., data = creditsub, method = "class")

# Display the results
rpart.plot(x = credit_model, yesno = 2, type = 0, extra = 0)
```

Train/test split
For this exercise, you'll randomly split the German Credit Dataset into two pieces: a training set (80%) called credit_train and a test set (20%) that we will call credit_test. We'll use these two sets throughout the chapter.

Instructions
100 XP
The credit data frame is loaded into the workspace.

Define n, the number of rows in the credit data frame.
Define n_train to be ~80% of n.
Set a seed (for reproducibility) and then sample n_train rows to define the set of training set indices.
Using row indices, subset the credit data frame to create two new datasets: credit_train and credit_test

```{r tts}
credit <- read.csv("credit.csv")

# Total number of rows in the credit data frame
n <- nrow(credit)

# Number of rows for the training set (80% of the dataset)
n_train <- round(.8 * n) 

# Create a vector of indices which is an 80% random sample
set.seed(123)
train_indices <- sample(1:n, n_train)

# Subset the credit data frame to training indices only
credit_train <- credit[train_indices, ]  
  
# Exclude the training indices to create the test set
credit_test <- credit[-train_indices, ] 
```

Train a classification tree model
In this exercise, you will train a model on the newly created training set and print the model object to get a sense of the results.

```{r tctm}
# Train the model (to predict 'default')
credit_model <- rpart(formula = default ~ ., data = credit_train, method = "class")

# Look at the model output                      
print(credit_model)
```

Compute confusion matrix
As discussed in the previous video, there are a number of different metrics by which you can measure the performance of a classification model. In this exercise, we will evaluate the performance of the model using test set classification error. A confusion matrix is a convenient way to examine the per-class error rates for all classes at once.

The confusionMatrix() function from the caret package prints both the confusion matrix and a number of other useful classification metrics such as "Accuracy" (fraction of correctly classified instances).

The caret package has been loaded for you.

Instructions
100 XP
Generate class predictions for the credit_test data frame using the credit_model object.
Using the caret::confusionMatrix() function, compute the confusion matrix for the test set.

```{r ccm}
# Generate predicted classes using the model object
class_prediction <- predict(object = credit_model,  
                        newdata = credit_test,   
                        type = "class")  
                            
# Calculate the confusion matrix for the test set
confusionMatrix(data = class_prediction,       
                reference = credit_test$default) 
```

